---
title: "Lake Trends"
author: "Anne Chambers"
date: "2025-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(cowplot)
library(lubridate)
theme_set(theme_cowplot())

source(here("analysis", "functions.R"))

# Import colors, specify names, etc.
lake_cols_raw <- lake_colors(raw_data = TRUE)
data_cols_warms <- data_colors(type = "warms")
data_cols <- data_colors(type = "assorted")
lake_names <- c("BIG HAWK LAKE", "LITTLE HAWK LAKE", "HALLS LAKE")
labs <- as_labeller(c("BIG HAWK LAKE" = "Big Hawk Lake", 
              "LITTLE HAWK LAKE" = "Little Hawk Lake", 
              "HALLS LAKE" = "Halls Lake"))
townships <- c("ALGONQUIN HIGHLANDS", "STANHOPE", "SHERBORNE",
               "MCCLINTOCK", "LIVINGSTONE", "NIGHTINGALE") # Lawrence not present in data
```

## Ice out dates

```{r}
# Import data
ice <- read_tsv(here("data", "ice_out.txt"))
lake_cols <- lake_colors()

# Adjust dates for plotting
ice$YMD <- lubridate::ymd(ice$YMD)
ice$doy <- yday(ice$YMD)

# Build plot
ice %>% 
  group_by(Year, Month) %>% 
  ggplot(aes(x = Year, y = doy, group = 1)) +
  geom_point(size = 3, color = lake_cols[["bighawk"]]) +
  geom_line(linewidth = 1, color = lake_cols[["bighawk"]]) +
  geom_smooth(method='lm',se=F, color = "grey40")+
  ylab("Day of the year") +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  # theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  ggtitle("Ice-out dates for Big Hawk Lake (1968-2025)")

lm(doy ~ Year, data = ice) %>% summary() # coefficient -0.08063 (for every 1 year change, the number of days earlier ice-out is happening)
# -0.08063*10 # every 10 years, ice out happening almost a year earlier
```

## Lake Health

These have been concatenated from various reports over the years, but most are available on the HHPLOA website [here](https://hallshawklakes.ca/environment/).

Trends both across entire system and within respective water bodies/river (e.g., is there a difference between Halls Lake and Big Hawk?)

### Map of monitoring stations

There are a number of stations on each lake at which measurements are taken.

```{r}
lat <- conv_unit(phos$`Latitude (DMS)`, from = "deg_min_sec", to = "dec_deg")
long <- conv_unit('104 38 26.88' , from = "deg_min_sec", to = "dec_deg")

phos %>% 
  dplyr::select(`Lake Name`, `Site ID`, `Latitude (DMS)`, `Long (DMS)`) %>% 
  distinct()
```

### Water transparency / clarity

Data obtained were obtained through the Lake Partner Program and are publicly available [here](https://data.ontario.ca/dataset/ontario-lake-partner/resource/3cd1cd13-5d7e-4bbb-b4c0-7d03508b41de).

```{r}
# Import the data
secchi_dat <- read_csv(here("data", "LPP_Secchi_2022.csv")) %>% 
  rename("Date" = `Date (DD-MMM-YY)`)
secchi_dat$Date <- lubridate::dmy(secchi_dat$Date)
secchi_dat <- secchi_dat %>% mutate("Year" = year(Date))

secchi_dat %>% 
  filter(TOWNSHIP %in% townships) %>% 
  # count(`Lake Name`) # 40 lakes in township
  filter(!`Lake Name` %in% lake_names) %>% 
  summarize(mean(`Secchi Depth (m)`)) # 4.85m (if including ours) or 4.62m if not including ours
```

Look at data for only our three lakes:

```{r}
secchi <- secchi_dat %>% 
  filter(TOWNSHIP == "STANHOPE") %>% 
  filter(`Lake Name` %in% lake_names)

secchi %>% summarize(mean(`Secchi Depth (m)`)) # 6.79m

# p_secchi <- 
  secchi %>% 
  ggplot(aes(x = Year, y = `Secchi Depth (m)`)) +
  # geom_point(aes(color = `Lake Name`)) +
  geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = labs, scales = "free") +
  scale_color_manual(values = lake_cols_raw) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank()) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(0, 11), breaks = 0:11) +
  ggtitle("Water transparency (1991-2022)") +
  ylab("Secchi depth (m)")

# Linear model for data points
```

Look at a single site for each lake, averaged over a year.

```{r}
secchi %>% 
  filter(`Lake Name` == "BIG HAWK LAKE" & `Site ID` == 3 |
           `Lake Name` == "LITTLE HAWK LAKE" & `Site ID` == 1 |
           `Lake Name` == "HALLS LAKE" & `Site ID` == 2) %>% 
  group_by(`Lake Name`, Year) %>% 
  summarize(mean_depth = mean(`Secchi Depth (m)`)) %>% 
  ggplot(aes(x = Year, y = mean_depth)) +
  geom_point(aes(color = `Lake Name`)) +
  geom_line(aes(color = `Lake Name`)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = labs, scales = "free") +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  scale_color_manual(values = lake_cols_raw) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Water transparency (1991-2022)") +
  ylab("Secchi depth (m)")

# # All in a single plot
# secchi %>% 
#   filter(`Lake Name` == "BIG HAWK LAKE" & `Site ID` == 3 |
#            `Lake Name` == "LITTLE HAWK LAKE" & `Site ID` == 1 |
#            `Lake Name` == "HALLS LAKE" & `Site ID` == 2) %>% 
#   group_by(`Lake Name`, Year) %>% 
#   summarize(mean_depth = mean(`Secchi Depth (m)`)) %>% 
#   ggplot(aes(x = Year, y = mean_depth, group = `Lake Name`)) +
#   geom_point(aes(color = `Lake Name`), size = 3) +
#   geom_line(aes(color = `Lake Name`), size = 1) +
#   geom_smooth(method = "lm", se = FALSE, 
#               aes(group = `Lake Name`, color = `Lake Name`), linetype = "dashed") +
#   scale_color_manual(values = lake_cols_raw, labels = labs) +
#   theme(strip.text = element_text(size = 16),
#         strip.background = element_blank()) +
#   ggtitle("Water transparency (1991-2022)") +
#   ylab("Secchi depth (m)")
```

Compare to two other monitored lakes in our township that have the most data:

```{r}
secchi_other <- read_csv(here("data", "LPP_Secchi_2022.csv")) %>% 
  filter(TOWNSHIP == "STANHOPE") %>% 
  filter(!`Lake Name` %in% lake_names) %>% 
  rename("Date" = `Date (DD-MMM-YY)`)
secchi_other$Date <- lubridate::dmy(secchi_other$Date)
secchi_other <- secchi_other %>% mutate("Year" = year(Date))

labs_other <- as_labeller(c("BOSHKUNG LAKE" = "Boshkung Lake", 
                            "KUSHOG LAKE" = "Kushog Lake"))

# Plot raw data
p_secchi_other <-
  secchi_other %>% 
  # Remove outlier point
  filter(`Secchi Depth (m)` < 15) %>% 
  filter(`Lake Name` == "BOSHKUNG LAKE" | `Lake Name` == "KUSHOG LAKE") %>% 
  ggplot(aes(x = Year, y = `Secchi Depth (m)`)) +
  # geom_point(aes(color = `Lake Name`)) +
  geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = labs_other, scales = "free") +
  scale_color_manual(values = c(data_cols_warms[[1]], data_cols_warms[[2]])) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank(),
        axis.title.y = element_blank()) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(0, 11), breaks = 0:11)

plot_grid(p_secchi + theme(axis.text.x = element_text(angle = 45, hjust = 1)), 
          p_secchi_other + theme(axis.text.x = element_text(angle = 45, hjust = 1)), 
          align = "h", axis = "t", nrow = 1, rel_widths = c(3, 2))
```

### Water quality

There are X metrics of water quality: pH levels, dissolved oxygen content, XXX

Phosphorus data were obtained through the Lake Partner Program and are publicly available [here](https://data.ontario.ca/dataset/ontario-lake-partner/resource/1a36234b-3b97-451c-8965-0050e7f283cf).

#### Phosphorus

```{r}
# Import data
phos_dat <- read_tsv(here("data", "LPP_Total_Phosphorus_2022.txt")) %>% 
    rename("Date" = `Date (DD-MMM-YY)`)
phos_dat$Date <- lubridate::dmy(phos_dat$Date)
phos_dat <- phos_dat %>% mutate("Year" = year(Date))
phos_dat %>% 
  filter(Township %in% townships) %>% 
  # filter(!`Lake Name` %in% lake_names) %>% 
  summarize(mean(`Average Total Phosphorus (µg/L)`)) # 6.43 (if including ours) or 6.52 if not including ours

phos <- phos_dat %>% 
  filter(`Lake Name` %in% lake_names)
# Mean per lake across all years
phos %>% summarize(mean(`Average Total Phosphorus (µg/L)`)) # 5.61

p_phos <-
  phos %>% 
  # Remove outliers (e.g., Big Hawk Lake in 2004)
  filter(`Possible outlier` == "No") %>% 
  ggplot(aes(x = Year, y = `Average Total Phosphorus (µg/L)`)) +
  geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
  # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = labs, scales = "free") +
  scale_color_manual(values = lake_cols_raw) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank()) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(0, 10), breaks = 0:10) +
  ggtitle("Phosphorus levels (2002-2022)") +
  ylab("Total phosphorus (µg/L)") +
  geom_hline(yintercept = 10, color = "darkred", linetype = "dashed")

# Compare to six other monitored lakes in Stanhope township:
phos_other <- phos_dat %>% 
  filter(Township == "STANHOPE") %>% 
  filter(!`Lake Name` %in% lake_names)

# colfunc <- colorRampPalette(c("#704c14", "#f07434", "#e0840c", "#bd724b"))

# Plot summary data, average per site on each lake
# phos_other %>% 
#   filter(`Possible outlier` == "No") %>% 
#   group_by(`Lake Name`, Year) %>% 
#   summarize(mean_phos = mean(`Average Total Phosphorus (µg/L)`)) %>% 
#   ggplot(aes(x = Year, y = mean_phos, group = `Lake Name`)) +
#   geom_point(aes(color = `Lake Name`)) +
#   geom_line(aes(color = `Lake Name`)) +
#   scale_y_continuous(limits = c(0, 11), breaks = 0:11) +
#   scale_color_manual(values = data_cols)

# Plot raw data
# p_phos_other <-
#   phos_other %>% 
#   filter(`Possible outlier` == "No") %>% 
#   ggplot(aes(x = Year, y = `Average Total Phosphorus (µg/L)`, group = `Lake Name`)) +
#   geom_point(aes(color = `Lake Name`)) +
#   geom_line(aes(color = `Lake Name`)) +
#   scale_y_continuous(limits = c(0, 18), breaks = 0:18) +
#   scale_color_manual(values = data_cols, labels = c("Big Brother Lake", "Boshkung Lake", 
#                                                     "Kabakwa Lake", "Kushog Lake", 
#                                                     "Little Cameron Lake", 
#                                                     "Saskatchewan Lake")) +
#   # scale_colour_manual(values = colfunc(6))
#   guides(color = guide_legend(position = "inside")) +
#   theme(legend.position.inside = c(0.35, 0.8),
#         legend.title = element_blank(),
#         legend.text = element_text(size = 10),
#         plot.title = element_text(face = "plain", size = 16, hjust = 0.5)) +
#   ggtitle("Other lakes in township")

  
# Plot only Boshkung Lake as a comparison
p_phos_bosh <-
  phos_other %>% 
  filter(`Possible outlier` == "No") %>% 
  filter(`Lake Name` == "BOSHKUNG LAKE") %>% 
  ggplot(aes(x = Year, y = `Average Total Phosphorus (µg/L)`, group = `Lake Name`)) +
  geom_point(aes(color = `Lake Name`), size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45") +
  # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  scale_y_continuous(limits = c(0, 18), breaks = 0:18) +
  guides(color = guide_legend(position = "inside")) +
  theme(legend.position = "none",
        plot.title = element_text(face = "plain", size = 16, hjust = 0.5)) +
  ggtitle("Boshkung Lake") +
  geom_hline(yintercept = 10, color = "darkred", linetype = "dashed")

plot_grid(p_phos, p_phos_bosh, align = "h", axis = "t", nrow = 1, rel_widths = c(3, 1))
```

Run stats:

```{r}
lm(doy ~ Year, data = ice) %>% summary() 
```

#### Calcium

```{r}
# Import data
calc_dat <- read_csv(here("data", "LPP_Calcium_2022.csv")) %>% 
    rename("Date" = `Date (DD-MMM-YY)`)
calc_dat$Date <- lubridate::dmy(calc_dat$Date)
calc_dat <- calc_dat %>% mutate("Year" = year(Date))
calc_dat %>% 
  filter(Township %in% townships) %>% 
  # filter(!`Lake Name` %in% lake_names) %>%
  summarize(mean(`Calcium (mg/L)`)) # 2.61 (if including ours) or 2.70 if not including ours

calc <- calc_dat %>% 
  filter(`Lake Name` %in% lake_names)
calc %>% summarize(mean(`Calcium (mg/L)`)) # 1.92
```

Build plot for Halls and Hawk Lakes.

```{r}
p_calc <-
  calc %>% 
  # It looks as though there's an outlier for Big Hawk Lake in 2004?
  ggplot(aes(x = Year, y = `Calcium (mg/L)`)) +
  geom_point(aes(color = `Lake Name`), size = 3) +
  # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  # geom_smooth(method = "lm", se = FALSE, color = "black", aes(group = 1)) +
  # geom_line(aes(color = `Lake Name`)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = labs, scales = "free") +
  scale_color_manual(values = lake_cols_raw) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank()) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(1, 3.1)) +
  ggtitle("Calcium levels (2008-2022)") +
  ylab("Calcium (mg/L)") +
  geom_hline(yintercept = 1.5, color = "darkred", linetype = "dashed")
```

Build plot for Boshkung Lake.

```{r}
# Compare to six other monitored lakes in Stanhope township:
calc_other <- calc_dat %>% 
  filter(Township == "STANHOPE") %>% 
  filter(!`Lake Name` %in% lake_names)

p_calc_kush <-
  calc_other %>% 
  filter(`Lake Name` == "KUSHOG LAKE") %>% 
  ggplot(aes(x = Year, y = `Calcium (mg/L)`)) +
  geom_point(aes(color = `Lake Name`), size = 3) +
  # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  # geom_smooth(method = "lm", se = FALSE, color = "black", aes(group = 1)) +
  # geom_line(aes(color = `Lake Name`)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  # facet_wrap(~`Lake Name`, labeller = labs, scales = "free") +
  # scale_color_manual(values = lake_cols_raw) +
  theme(legend.position = "none",
        plot.title = element_text(face = "plain", size = 16, hjust = 0.5)) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(1, 3.1)) +
  ggtitle("Kushog Lake") +
  ylab("Calcium (mg/L)") +
  geom_hline(yintercept = 1.5, color = "darkred", linetype = "dashed")
```

Panel both plots together.

```{r}
plot_grid(p_calc, p_calc_kush, align = "h", axis = "t", nrow = 1, rel_widths = c(3, 1))
```

#### Dissolved O2

Import and tidy data.

```{r}
lake_col_full <- lake_colors(full_names = TRUE)

halls_disso <- read_tsv(here("data", "halls_water_chemistry.txt")) %>% 
  pivot_longer(cols = 4:20, names_to = "site_code", values_to = "value") %>% 
  # Average between replicates
  group_by(Year, site_code, Parameter) %>% 
  mutate(avg_value = mean(value),
         site_code_norep = site_code)
hawk_disso <- read_tsv(here("data", "hawk_water_chemistry.txt")) %>% 
  pivot_longer(cols = 4:19, names_to = "site_code", values_to = "value") %>% 
  # Average between replicates
  group_by(Year, site_code, Parameter) %>% 
  mutate(avg_value = mean(value),
         site_code_norep = site_code)
# Join datasets together
disso <- bind_rows(halls_disso, hawk_disso)

# Remove replicates
disso$site_code_norep <- gsub('-R1', '', disso$site_code_norep)
disso$site_code_norep <- gsub('-R2', '', disso$site_code_norep)
disso <- disso %>% 
  ungroup() %>% 
  dplyr::select(-c(site_code, value)) %>% 
  distinct()

# Add in proper lake names
disso <-
  disso %>% 
  mutate(`Lake Name` = case_when(str_detect(site_code_norep, regex("HALL")) ~ "Halls Lake",
                                 str_detect(site_code_norep, regex("BHWK")) ~ "Big Hawk Lake",
                                 str_detect(site_code_norep, regex("LHWK")) ~ "Little Hawk Lake"))
```

Plot dissolved oxygen.

```{r}
disso %>% 
  filter(Parameter == "Diss_O2") %>% 
  ggplot(aes(x = Year, y = avg_value)) +
  geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
  # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, scales = "free") +
  scale_color_manual(values = lake_col_full) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank()) +
  # scale_x_continuous(breaks = scales::breaks_width(5)) +
  # scale_y_continuous(limits = c(0, 10), breaks = 0:10) +
  ggtitle("Dissolved oxygen levels (2019-2024)") +
  ylab("Dissolved O2 (mg/L)")
```

### Benthic monitoring

Import and tidy data.

```{r}
halls_indices <- read_tsv(here("data", "halls_indices.txt")) %>% 
  pivot_longer(cols = 4:24, names_to = "site_code", values_to = "value") %>% 
  # Average between replicates
  group_by(Year, site_code, Biotic_index) %>% 
  mutate(avg_value = mean(value),
         site_code_norep = site_code)
hawk_indices <- read_tsv(here("data", "hawk_indices.txt")) %>% 
  pivot_longer(cols = 4:19, names_to = "site_code", values_to = "value") %>% 
  # Average between replicates
  group_by(Year, site_code, Biotic_index) %>% 
  mutate(avg_value = mean(value),
         site_code_norep = site_code)
# Join datasets together
indices <- bind_rows(halls_indices, hawk_indices)

# Remove replicates
indices$site_code_norep <- gsub('-R1', '', indices$site_code_norep)
indices$site_code_norep <- gsub('-R2', '', indices$site_code_norep)
indices <- indices %>% 
  ungroup() %>% 
  dplyr::select(-c(site_code, value)) %>% 
  distinct()

# Add in proper lake names
indices <-
  indices %>% 
  mutate(`Lake Name` = case_when(str_detect(site_code_norep, regex("HALL")) ~ "Halls Lake",
                                 str_detect(site_code_norep, regex("BHWK")) ~ "Big Hawk Lake",
                                 str_detect(site_code_norep, regex("LHWK")) ~ "Little Hawk Lake"))

rect_data <- read_tsv(here("data", "index_rect_data.txt"))
```

#### Diversity (Simpson's D)

```{r}
indices %>% 
  filter(Biotic_index == "Simpsons_Index") %>% 
  ggplot(aes(x = Year, y = avg_value)) +
  geom_rect(data = rect_data %>% filter(Biotic_index == "Simpsons_Index"), inherit.aes = FALSE, 
            aes(xmin=-Inf, xmax=+Inf, ymin = ymin, ymax = ymax, fill = Category), alpha = 0.4) +
  scale_fill_manual(values = c("Poor" = "#d53e4f", "Fair" = "#fee08b", "Good" = "#abdda4", "Excellent" = "#3288bd")) +
  geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
  geom_point(aes(group = Year), size = 3, pch = 21, color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, scales = "free") +
  scale_color_manual(values = lake_col_full) +
  theme(legend.position = "none",
      strip.text = element_text(size = 16),
      strip.background = element_blank()) +
  scale_y_continuous(limits = c(0, 1)) +
  ggtitle("Diversity (2020-2024)") +
  ylab("Simpson's diversity index")
```

#### Tolerance (mHBI)

```{r}
indices %>% 
  filter(Biotic_index == "mHBI") %>% 
  ggplot(aes(x = Year, y = avg_value)) +
  geom_rect(data = rect_data %>% filter(Biotic_index == "mHBI"), inherit.aes = FALSE,
  aes(xmin=-Inf, xmax=+Inf, ymin = ymin, ymax = ymax, fill = Category), alpha = 0.4) +
  scale_fill_manual(values = c("Poor" = "#d53e4f", "Fair" = "#fee08b", "Good" = "#abdda4", "Excellent" = "#3288bd")) +
  geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
  geom_point(aes(group = Year), size = 3, pch = 21, color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, scales = "free") +
  scale_color_manual(values = lake_col_full) +
  theme(legend.position = "none",
      strip.text = element_text(size = 16),
      strip.background = element_blank()) +
  scale_y_continuous(limits = c(0, 10)) +
  ggtitle("Pollution tolerance (2020-2024)") +
  ylab("Modified HBI (family level)")
"Very poor" = "#9e0142"
"Poor" = "#d53e4f"
"Fairly poor" = "#fdae61"
"Fair" = "#fde18b"
"Good" = "#e6f598"
"Very good" = "#abdda4"
"Excellent" = "#3288bd"
```

#### Percent EOT

```{r}
indices %>% 
  filter(Biotic_index == "Perc_EOT") %>% 
  ggplot(aes(x = Year, y = avg_value)) +
  geom_rect(data = rect_data %>% filter(Biotic_index == "Perc_EOT"), inherit.aes = FALSE,
  aes(xmin=-Inf, xmax=+Inf, ymin = ymin, ymax = ymax, fill = Category), alpha = 0.4) +
  scale_fill_manual(values = c("Fair" = "#fdae61", "Good" = "#abdda4", "Excellent" = "#3288bd")) +
  geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
  geom_point(aes(group = Year), size = 3, pch = 21, color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, scales = "free") +
  scale_color_manual(values = lake_col_full) +
  theme(legend.position = "none",
      strip.text = element_text(size = 16),
      strip.background = element_blank()) +
  ggtitle("Diversity of pollution-intolerant groups (2020-2024)") +
  ylab("Percent EOT")
```

## Maps

5)  Nice data viz of Trent-Severn waterway
6)  Possibly also a higher resolution map of our watershed (the one in the 2006 summary is really low resolution) Based on results from analyses above, I will also fact-check any claims made in the summary (e.g., statements about the acidification of Little Hawk, etc.)

## Other analyses

1) Algae growth 2) Bacterial levels 3) If a trend is found with ice out dates, look into finding data 4) Lake health comparison to other lakes in the area
