---
title: "Lake Trends"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(cowplot)
library(lubridate)
library(grid)
library(gridExtra)
library(tidyterra)
theme_set(theme_cowplot())

source(here("analysis", "functions.R"))

# Import colors, specify names, etc.
lake_cols_raw <- lake_colors(raw_data = TRUE)
data_cols_warms <- data_colors(type = "warms")
data_cols <- data_colors(type = "assorted")
lake_col_full <- lake_colors(full_names = TRUE)
lake_names <- c("BIG HAWK LAKE", "LITTLE HAWK LAKE", "HALLS LAKE")
labs <- as_labeller(c("BIG HAWK LAKE" = "Big Hawk Lake", 
              "LITTLE HAWK LAKE" = "Little Hawk Lake", 
              "HALLS LAKE" = "Halls Lake"))
townships <- c("ALGONQUIN HIGHLANDS", "STANHOPE", "SHERBORNE",
               "MCCLINTOCK", "LIVINGSTONE", "NIGHTINGALE") # Lawrence not present in data
```

## Ice out dates

```{r}
# Import data
ice <- read_tsv(here("data", "ice_out.txt"))
lake_cols <- lake_colors()

# Adjust dates for plotting
ice$YMD <- lubridate::ymd(ice$YMD)
ice$doy <- yday(ice$YMD)
```

Build plot:

```{r}
p_ice <- ice %>% 
  group_by(Year, Month) %>% 
  ggplot(aes(x = Year, y = doy, group = 1)) +
  # geom_point(size = 3, color = lake_cols[["bighawk"]]) +
  # geom_line(linewidth = 1, color = lake_cols[["bighawk"]]) +
  geom_point(size = 2, color = "#b4d6e7") +
  geom_line(linewidth = 1, color = "#b4d6e7") +
  geom_smooth(method = 'lm', se = F, color = "#186484") +
  # ylab("Day of the year") +
  ylab("Ice out date") +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  # ggtitle("Ice-out dates for Big Hawk Lake (1968-2025)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        axis.title = element_text(size = 12))

# Save plot in "plots" directory
ggsave(here("plots", "ice_out.pdf"), width = 3.5, height = 2.5)

# For AGM pres
ice %>% 
  group_by(Year, Month) %>% 
  ggplot(aes(x = Year, y = doy, group = 1)) +
  geom_point(size = 4, color = "#b4d6e7") +
  geom_line(linewidth = 2, color = "#b4d6e7") +
  geom_smooth(method = 'lm', se = F, color = "#186484") +
  ylab("Ice out date") +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  ggtitle("Ice-out dates for Big Hawk Lake (1968-2025)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 16)) +
    geom_point(data = ice %>% filter(Year == 2012 | Year == 1992), 
             aes(x = Year, y = doy), size = 4, pch = 21, color = "black")

ggsave(here("plots", "ice_out_pres.pdf"), width = 8, height = 5.5)
```

Find earliest and latest ice-out dates:

```{r}
ice %>% filter(doy == min(doy)) # earliest was on March 26, 2012
ice %>% filter(doy == max(doy)) # latest was on May 8, 1992
```

Run stats on data:

```{r}
lm(doy ~ Year, data = ice) %>% summary() # coefficient -0.08063 (for every 1 year change, the number of days earlier ice-out is happening)
# -0.08063*10 # every 10 years, ice out happening almost a year earlier
```

## Lake Health

These have been concatenated from various reports over the years, but most are available on the HHPLOA website [here](https://hallshawklakes.ca/environment/).

Trends both across entire system and within respective water bodies/river (e.g., is there a difference between Halls Lake and Big Hawk?)

### Water transparency / clarity

Data obtained were obtained through the Lake Partner Program and are publicly available [here](https://data.ontario.ca/dataset/ontario-lake-partner/resource/3cd1cd13-5d7e-4bbb-b4c0-7d03508b41de).

```{r}
# Import the data
secchi_dat <- read_csv(here("data", "LPP_Secchi_2022.csv")) %>% 
  rename("Date" = `Date (DD-MMM-YY)`)
secchi_dat$Date <- lubridate::dmy(secchi_dat$Date)
secchi_dat <- secchi_dat %>% mutate("Year" = year(Date))

secchi_dat %>% 
  filter(TOWNSHIP %in% townships) %>% 
  # count(`Lake Name`) # 40 lakes in township
  filter(!`Lake Name` %in% lake_names) %>% 
  summarize(mean(`Secchi Depth (m)`)) # 4.85m (if including ours) or 4.62m if not including ours
```

Look at data for only our three lakes:

```{r}
secchi <- secchi_dat %>% 
  filter(TOWNSHIP == "STANHOPE") %>% 
  filter(`Lake Name` %in% lake_names)

secchi %>% summarize(mean(`Secchi Depth (m)`)) # 6.79m

p_secchi <-
  secchi %>% 
  ggplot(aes(x = Year, y = `Secchi Depth (m)`)) +
  geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = labs, scales = "free") +
  scale_color_manual(values = lake_cols_raw) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank()) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(0, 11), breaks = 0:11) +
  ggtitle("Water transparency (1991-2022)") +
  ylab("Secchi depth (m)")

# Save plot in "plots" directory
p_secchi
ggsave(here("plots", "secchi_hhl.pdf"), width = 12, height = 6)
```

Compare to two other monitored lakes in our township that have the most data:

```{r}
secchi_other <- read_csv(here("data", "LPP_Secchi_2022.csv")) %>% 
  # filter(TOWNSHIP == "STANHOPE") %>% 
  filter(!`Lake Name` %in% lake_names) %>% 
  rename("Date" = `Date (DD-MMM-YY)`)
secchi_other$Date <- lubridate::dmy(secchi_other$Date)
secchi_other <- secchi_other %>% mutate("Year" = year(Date))

# Plot raw data
p_secchi_other <-
  secchi_other %>% 
  # Remove outlier point
  filter(`Secchi Depth (m)` < 15) %>% 
  filter(`Lake Name` == "KENNISIS LAKE" | `Lake Name` == "KUSHOG LAKE") %>% 
  ggplot(aes(x = Year, y = `Secchi Depth (m)`)) +
  # geom_point(aes(color = `Lake Name`)) +
  geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = as_labeller(c("KENNISIS LAKE" = "Kennisis Lake", 
                                                    "KUSHOG LAKE" = "Kushog Lake")), 
             scales = "free") +
  scale_color_manual(values = c("KUSHOG LAKE" = data_cols_warms[[2]], 
                                "KENNISIS LAKE" = data_cols_warms[[3]])) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank(),
        axis.title.y = element_blank()) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(0, 11), breaks = 0:11)

plot_grid(p_secchi + theme(axis.text.x = element_text(angle = 45, hjust = 1)), 
          p_secchi_other + theme(axis.text.x = element_text(angle = 45, hjust = 1)), 
          align = "h", axis = "t", nrow = 1, rel_widths = c(3, 2))
ggsave(here("plots", "secchi_all.pdf"), width = 14, height = 6)
```

Run stats.

```{r}
# Linear model for data points
lm(`Secchi Depth (m)` ~ Year, data = secchi %>% filter(`Lake Name` == "BIG HAWK LAKE")) %>% summary()
lm(`Secchi Depth (m)` ~ Year, data = secchi %>% filter(`Lake Name` == "LITTLE HAWK LAKE")) %>% summary()
lm(`Secchi Depth (m)` ~ Year, data = secchi %>% filter(`Lake Name` == "HALLS LAKE")) %>% summary()

# lm(`Secchi Depth (m)` ~ Year, data = secchi_dat %>% filter(`Lake Name` == "BOSHKUNG LAKE")) %>% summary()
lm(`Secchi Depth (m)` ~ Year, data = secchi_dat %>% filter(`Lake Name` == "KENNISIS LAKE")) %>% summary()
lm(`Secchi Depth (m)` ~ Year, data = secchi_dat %>% filter(`Lake Name` == "KUSHOG LAKE")) %>% summary()
```

Calculate average clarity in each lake for each year.

```{r}
sum <- secchi %>% 
  # filter(`Lake Name` == "HALLS LAKE") %>% 
  group_by(Year) %>% 
  summarize(mean_per_year = mean(`Secchi Depth (m)`))

# Average decline in clarity across all three lakes
secchi %>% 
  filter(`Lake Name` == "HALLS LAKE") %>% 
  filter(Year == 1991 | Year == 2022) %>% 
  group_by(Year) %>% 
  summarize(mean_per_year = mean(`Secchi Depth (m)`))
```

### Water quality

There are three metrics of water quality: total phosphorus, calcium, and dissolved oxygen.

Phosphorus data were obtained through the Lake Partner Program and are publicly available [here](https://data.ontario.ca/dataset/ontario-lake-partner/resource/1a36234b-3b97-451c-8965-0050e7f283cf).

#### Phosphorus

```{r}
# Import data
phos_dat <- read_csv(here("data", "LPP_Total_Phosphorus_2022.csv")) %>% 
    rename("Date" = `Date (DD-MMM-YY)`)
phos_dat$Date <- lubridate::dmy(phos_dat$Date)
phos_dat <- phos_dat %>% mutate("Year" = year(Date))
phos_dat %>% 
  filter(Township %in% townships) %>% 
  # filter(!`Lake Name` %in% lake_names) %>% 
  summarize(mean(`Average Total Phosphorus (µg/L)`)) # 6.43 (if including ours) or 6.52 if not including ours

phos <- phos_dat %>% 
  filter(`Lake Name` %in% lake_names)
# Mean per lake across all years
phos %>% summarize(mean(`Average Total Phosphorus (µg/L)`)) # 5.61

# Compare to six other monitored lakes in Stanhope township:
phos_other <- phos_dat %>% 
  filter(Township == "STANHOPE") %>% 
  filter(!`Lake Name` %in% lake_names)
```

Build plots: 

```{r}
p_phos <-
  phos %>% 
  # Remove outliers (e.g., Big Hawk Lake in 2004)
  filter(`Possible outlier` == "No") %>% 
  ggplot(aes(x = Year, y = `Average Total Phosphorus (µg/L)`)) +
  geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
  # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = labs, scales = "free") +
  scale_color_manual(values = lake_cols_raw) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank(),
        axis.title.x = element_blank()) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(0, 10), breaks = 0:10) +
  ggtitle("Phosphorus levels (2002-2022)") +
  ylab("Total phosphorus (µg/L)") +
  geom_hline(yintercept = 10, color = "darkred", linetype = "dashed")

p_phos
ggsave(here("plots", "phos_hhl.pdf"), width = 12, height = 6)

# Plot Kushog and Kennisis Lakes as a comparison
p_phos_kk <-
  phos_dat %>% 
  filter(`Possible outlier` == "No") %>% 
  filter(`Lake Name` == "KUSHOG LAKE" | `Lake Name` == "KENNISIS LAKE") %>% 
  ggplot(aes(x = Year, y = `Average Total Phosphorus (µg/L)`)) +
  geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
  # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = as_labeller(c("KENNISIS LAKE" = "Kennisis Lake", 
                                                    "KUSHOG LAKE" = "Kushog Lake")), 
             scales = "free") +
  scale_color_manual(values = c("KUSHOG LAKE" = "#f07434", "KENNISIS LAKE" = "#a82c94")) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank(),
        axis.title = element_blank()) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(0, 12), breaks = 0:12) +
  ylab("Total phosphorus (µg/L)") +
  geom_hline(yintercept = 10, color = "darkred", linetype = "dashed")
  
p_tp <- plot_grid(p_phos, p_phos_kk, align = "h", axis = "t", nrow = 1, rel_widths = c(3, 2))

# Add common x axis to plot
x.grob <- textGrob("Year", gp = gpar(col = "black", fontsize = 16))
grid.arrange(arrangeGrob(p_tp, bottom = x.grob))

# Save plot
ggsave(here("plots", "phos_all.pdf"), width = 14, height = 6)
```

Plot only a single other lake:

```{r}
# Plot only one other lake as a comparison (Kushog below)
phos_dat %>% 
  filter(`Possible outlier` == "No") %>% 
  filter(`Lake Name` == "KUSHOG LAKE") %>% 
  ggplot(aes(x = Year, y = `Average Total Phosphorus (µg/L)`, group = `Lake Name`)) +
  geom_point(color = "#f07434", size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45") +
  # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  scale_y_continuous(limits = c(0, 10), breaks = 0:10) +
  guides(color = guide_legend(position = "inside")) +
  theme(legend.position = "none",
        plot.title = element_text(face = "plain", size = 16, hjust = 0.5),
        axis.title.x = element_blank()) +
  ggtitle("Kushog Lake") +
  geom_hline(yintercept = 10, color = "darkred", linetype = "dashed")
```

Plot averages per site on each lake (other than our three):

```{r}
# Plot summary data, average per site on each lake
phos_other %>%
  filter(`Possible outlier` == "No") %>%
  group_by(`Lake Name`, Year) %>%
  summarize(mean_phos = mean(`Average Total Phosphorus (µg/L)`)) %>%
  ggplot(aes(x = Year, y = mean_phos, group = `Lake Name`)) +
  geom_point(aes(color = `Lake Name`)) +
  geom_line(aes(color = `Lake Name`)) +
  scale_y_continuous(limits = c(0, 11), breaks = 0:11) +
  scale_color_manual(values = data_cols)
```

Run stats:

```{r}
lm(`Average Total Phosphorus (µg/L)` ~ Year, data = phos %>% 
  filter(`Lake Name` == "BIG HAWK LAKE")) %>%
  summary()

lm(`Average Total Phosphorus (µg/L)` ~ Year, data = phos %>% 
  filter(`Lake Name` == "LITTLE HAWK LAKE")) %>%
  summary()

lm(`Average Total Phosphorus (µg/L)` ~ Year, data = phos %>% 
  filter(`Lake Name` == "HALLS LAKE")) %>%
  summary()

lm(`Average Total Phosphorus (µg/L)` ~ Year, data = phos_dat %>% 
  filter(`Lake Name` == "KENNISIS LAKE")) %>%
  summary()

lm(`Average Total Phosphorus (µg/L)` ~ Year, data = phos_dat %>% 
  filter(`Lake Name` == "KUSHOG LAKE")) %>%
  summary()
```

Do highest values tend to be in particular spots around the lake?

```{r}
p_sep_bhawk <-
  phos %>% 
  filter(`Lake Name` == "BIG HAWK LAKE") %>% 
  filter(`Possible outlier` == "No") %>% 
  ggplot(aes(x = Year, y = `Average Total Phosphorus (µg/L)`)) +
  geom_point(aes(color = `Site Description`), size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45") +
  # scale_y_continuous(limits = c(0, 18), breaks = 0:18) +
  theme(plot.title = element_text(face = "plain", size = 16, hjust = 0.5),
        legend.position.inside = c(0.4, 0.85)) +
  ggtitle("Big Hawk Lake") +
  guides(color = guide_legend(position = "inside")) +
  geom_hline(yintercept = 10, color = "darkred", linetype = "dashed")

p_sep_lhawk <-
  phos %>% 
  filter(`Lake Name` == "LITTLE HAWK LAKE") %>% 
  filter(`Possible outlier` == "No") %>% 
  ggplot(aes(x = Year, y = `Average Total Phosphorus (µg/L)`)) +
  geom_point(aes(color = `Site Description`), size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45") +
  # scale_y_continuous(limits = c(0, 18), breaks = 0:18) +
  theme(plot.title = element_text(face = "plain", size = 16, hjust = 0.5),
        legend.position.inside = c(0.4, 0.85)) +
  ggtitle("Little Hawk Lake") +
  guides(color = guide_legend(position = "inside")) +
  geom_hline(yintercept = 10, color = "darkred", linetype = "dashed")

p_sep_halls <-
  phos %>% 
  filter(`Lake Name` == "HALLS LAKE") %>% 
  filter(`Possible outlier` == "No") %>% 
  ggplot(aes(x = Year, y = `Average Total Phosphorus (µg/L)`)) +
  geom_point(aes(color = `Site Description`), size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45") +
  # scale_y_continuous(limits = c(0, 18), breaks = 0:18) +
  theme(plot.title = element_text(face = "plain", size = 16, hjust = 0.5),
        legend.position.inside = c(0.4, 0.85)) +
  ggtitle("Halls Lake") +
  guides(color = guide_legend(position = "inside")) +
  geom_hline(yintercept = 10, color = "darkred", linetype = "dashed")

plot_grid(p_sep_bhawk, p_sep_lhawk, p_sep_halls, nrow = 1)
```

#### Calcium

```{r}
# Import data
calc_dat <- read_csv(here("data", "LPP_Calcium_2022.csv")) %>% 
    rename("Date" = `Date (DD-MMM-YY)`)
calc_dat$Date <- lubridate::dmy(calc_dat$Date)
calc_dat <- calc_dat %>% mutate("Year" = year(Date))
calc_dat %>% 
  filter(Township %in% townships) %>% 
  # filter(!`Lake Name` %in% lake_names) %>%
  summarize(mean(`Calcium (mg/L)`)) # 2.61 (if including ours) or 2.70 if not including ours

calc <- calc_dat %>% 
  filter(`Lake Name` %in% lake_names)
calc %>% summarize(mean(`Calcium (mg/L)`)) # 1.92
```

Build plot for Halls and Hawk Lakes.

```{r}
p_calc <-
  calc %>% 
  # It looks as though there's an outlier for Big Hawk Lake in 2004?
  ggplot(aes(x = Year, y = `Calcium (mg/L)`)) +
  geom_point(aes(color = `Lake Name`), size = 3) +
  # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  # geom_smooth(method = "lm", se = FALSE, color = "black", aes(group = 1)) +
  # geom_line(aes(color = `Lake Name`)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = labs, scales = "free") +
  scale_color_manual(values = lake_cols_raw) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank(),
        axis.title.x = element_blank()) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(1, 3.1)) +
  ggtitle("Calcium levels (2008-2022)") +
  ylab("Calcium (mg/L)") +
  geom_hline(yintercept = 1.5, color = "darkred", linetype = "dashed")

# Save plot
p_calc
ggsave(here("plots", "calc_hhl.pdf"), width = 12, height = 6)

```

Build plot for Kushog and Kennisis Lakes.

```{r}
# Compare to other monitored lakes in Stanhope township:
calc_other <- calc_dat %>% 
  filter(Township == "STANHOPE") %>% 
  filter(!`Lake Name` %in% lake_names)

p_calc_kk <-
  calc_dat %>% 
  filter(`Lake Name` == "KUSHOG LAKE" | `Lake Name` == "KENNISIS LAKE") %>% 
  ggplot(aes(x = Year, y = `Calcium (mg/L)`)) +
  geom_point(aes(color = `Lake Name`), size = 3) +
  # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
  # geom_smooth(method = "lm", se = FALSE, color = "black", aes(group = 1)) +
  # geom_line(aes(color = `Lake Name`)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, labeller = as_labeller(c("KENNISIS LAKE" = "Kennisis Lake", "KUSHOG LAKE" = "Kushog Lake")), scales = "free") +
  scale_color_manual(values = c("KUSHOG LAKE" = "#f07434", "KENNISIS LAKE" = "#a82c94")) +
  theme(legend.position = "none",
        strip.text = element_text(size = 16),
        strip.background = element_blank(),
        axis.title = element_blank()) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  scale_y_continuous(limits = c(1, 3.1)) +
  ylab("Calcium (mg/L)") +
  geom_hline(yintercept = 1.5, color = "darkred", linetype = "dashed")
```

Panel both plots together and save.

```{r}
p_c <- plot_grid(p_calc, p_calc_kk, align = "h", axis = "t", nrow = 1, rel_widths = c(3, 2))
grid.arrange(arrangeGrob(p_c, bottom = x.grob))

# Save plot
ggsave(here("plots", "calc_all.pdf"), width = 14, height = 6)
```

Run stats:

```{r}
lm(`Calcium (mg/L)` ~ Year, data = calc %>% 
  filter(`Lake Name` == "BIG HAWK LAKE")) %>%
  summary()
lm(`Calcium (mg/L)` ~ Year, data = calc %>% 
  filter(`Lake Name` == "HALLS LAKE")) %>%
  summary()
lm(`Calcium (mg/L)` ~ Year, data = calc %>% 
  filter(`Lake Name` == "LITTLE HAWK LAKE")) %>%
  summary()

lm(`Calcium (mg/L)` ~ Year, data = calc_dat %>% 
  filter(`Lake Name` == "KENNISIS LAKE")) %>%
  summary()
lm(`Calcium (mg/L)` ~ Year, data = calc_dat %>% 
  filter(`Lake Name` == "KUSHOG LAKE")) %>%
  summary()
```

#### Phosphorus & calcium: same plot

```{r}
p_tp_comb <- plot_grid(p_phos + ggtitle("Phosphorus and calcium levels (2002-2022)"), 
                  p_phos_kk, align = "h", axis = "t", nrow = 1, rel_widths = c(3, 2))

p_c_comb <- plot_grid(p_calc + scale_x_continuous(limits = c(2002, 2022)) + ggtitle(""), 
                      p_calc_kk + scale_x_continuous(limits = c(2002, 2022)), 
                      align = "h", axis = "t", nrow = 1, rel_widths = c(3, 2))

p_comb <- plot_grid(p_tp_comb, p_c_comb, align = "h", axis = "t", nrow = 2)
grid.arrange(arrangeGrob(p_comb, bottom = x.grob))
ggsave(here("plots", "phoscalc_all.pdf"), width = 14, height = 8)
```

#### Dissolved O2

Import and tidy data.

```{r}
halls_disso <- read_tsv(here("data", "halls_water_chemistry.txt")) %>% 
  pivot_longer(cols = 4:20, names_to = "site_code", values_to = "value") %>% 
  # Average between replicates
  group_by(Year, site_code, Parameter) %>% 
  mutate(avg_value = mean(value),
         site_code_norep = site_code)
hawk_disso <- read_tsv(here("data", "hawk_water_chemistry.txt")) %>% 
  pivot_longer(cols = 4:19, names_to = "site_code", values_to = "value") %>% 
  # Average between replicates
  group_by(Year, site_code, Parameter) %>% 
  mutate(avg_value = mean(value),
         site_code_norep = site_code)
# Join datasets together
disso <- bind_rows(halls_disso, hawk_disso)

# Remove replicates
disso$site_code_norep <- gsub('-R1', '', disso$site_code_norep)
disso$site_code_norep <- gsub('-R2', '', disso$site_code_norep)
disso <- disso %>% 
  ungroup() %>% 
  dplyr::select(-c(site_code, value)) %>% 
  distinct()

# Add in proper lake names
disso <-
  disso %>% 
  mutate(`Lake Name` = case_when(str_detect(site_code_norep, regex("HALL")) ~ "Halls Lake",
                                 str_detect(site_code_norep, regex("BHWK")) ~ "Big Hawk Lake",
                                 str_detect(site_code_norep, regex("LHWK")) ~ "Little Hawk Lake"))
```

We decided to not included the plotted dissolved oxygen as the data are so limited. But if you want to, uncomment below and run:

```{r}
# disso %>% 
#   filter(Parameter == "Diss_O2") %>% 
#   ggplot(aes(x = Year, y = avg_value)) +
#   geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
#   # geom_boxplot(aes(color = `Lake Name`, group = Year)) +
#   geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
#   facet_wrap(~`Lake Name`, scales = "free") +
#   scale_color_manual(values = lake_col_full) +
#   theme(legend.position = "none",
#         strip.text = element_text(size = 16),
#         strip.background = element_blank()) +
#   # scale_x_continuous(breaks = scales::breaks_width(5)) +
#   # scale_y_continuous(limits = c(0, 10), breaks = 0:10) +
#   ggtitle("Dissolved oxygen levels (2019-2024)") +
#   ylab("Dissolved O2 (mg/L)")
# 
# # Save plot
# ggsave(here("plots", "disso_hhl.pdf"), width = 12, height = 8)
```

### Benthic monitoring

Import and tidy data.

```{r}
halls_indices <- read_tsv(here("data", "halls_indices.txt")) %>% 
  pivot_longer(cols = 4:24, names_to = "site_code", values_to = "value") %>% 
  # Average between replicates
  group_by(Year, site_code, Biotic_index) %>% 
  mutate(avg_value = mean(value),
         site_code_norep = site_code)
hawk_indices <- read_tsv(here("data", "hawk_indices.txt")) %>% 
  pivot_longer(cols = 4:19, names_to = "site_code", values_to = "value") %>% 
  # Average between replicates
  group_by(Year, site_code, Biotic_index) %>% 
  mutate(avg_value = mean(value),
         site_code_norep = site_code)
# Join datasets together
indices <- bind_rows(halls_indices, hawk_indices)

# Remove replicates
indices$site_code_norep <- gsub('-R1', '', indices$site_code_norep)
indices$site_code_norep <- gsub('-R2', '', indices$site_code_norep)
indices <- indices %>% 
  ungroup() %>% 
  dplyr::select(-c(site_code, value)) %>% 
  distinct()

# Add in proper lake names
indices <-
  indices %>% 
  mutate(`Lake Name` = case_when(str_detect(site_code_norep, regex("HALL")) ~ "Halls Lake",
                                 str_detect(site_code_norep, regex("BHWK")) ~ "Big Hawk Lake",
                                 str_detect(site_code_norep, regex("LHWK")) ~ "Little Hawk Lake"))

# Ranges for tolerances for each metric
rect_data <- read_tsv(here("data", "index_rect_data.txt"))
```

#### Diversity (Simpson's D)

```{r}
indices %>% 
  filter(Biotic_index == "Simpsons_Index") %>% 
  ggplot(aes(x = Year, y = avg_value)) +
  geom_rect(data = rect_data %>% filter(Biotic_index == "Simpsons_Index"), inherit.aes = FALSE, 
            aes(xmin = -Inf, xmax = +Inf, ymin = ymin, ymax = ymax, fill = Category), alpha = 0.4) +
  scale_fill_manual(values = c("Poor" = "#d53e4f", "Fair" = "#fee08b", "Good" = "#abdda4", "Excellent" = "#3288bd")) +
  geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
  geom_point(aes(group = Year), size = 3, pch = 21, color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, scales = "free") +
  scale_color_manual(values = lake_col_full) +
  theme(legend.position = "none",
      strip.text = element_text(size = 16),
      strip.background = element_blank()) +
  scale_y_continuous(limits = c(0, 1)) +
  ggtitle("Diversity (2020-2024)") +
  ylab("Simpson's diversity index")

# Save plot
ggsave(here("plots", "indices_hhl.pdf"), width = 12, height = 8)
```

Run stats:

```{r}
lm(avg_value ~ Year, 
   data = indices %>% 
     filter(Biotic_index == "Simpsons_Index") %>% 
     filter(`Lake Name` == "Big Hawk Lake")) %>%
  summary() # coeff 0.02141, non sig
lm(avg_value ~ Year, 
   data = indices %>% 
     filter(Biotic_index == "Simpsons_Index") %>% 
     filter(`Lake Name` == "Halls Lake")) %>%
  summary() # coeff -0.01717, non sig
lm(avg_value ~ Year, 
   data = indices %>% 
     filter(Biotic_index == "Simpsons_Index") %>% 
     filter(`Lake Name` == "Little Hawk Lake")) %>%
  summary() # coeff 0.0006307, non sig
```

#### Tolerance (mHBI)

```{r}
indices %>% 
  filter(Biotic_index == "mHBI") %>% 
  ggplot(aes(x = Year, y = avg_value)) +
  geom_rect(data = rect_data %>% filter(Biotic_index == "mHBI"), inherit.aes = FALSE,
  aes(xmin=-Inf, xmax=+Inf, ymin = ymin, ymax = ymax, fill = Category), alpha = 0.4) +
  scale_fill_manual(values = c("Very poor" = "#9e0142", "Poor" = "#d53e4f", "Fairly poor" = "#fdae61", "Fair" = "#fde18b", "Good" = "#e6f598", "Very good" = "#abdda4", "Excellent" = "#3288bd")) +
  geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
  geom_point(aes(group = Year), size = 3, pch = 21, color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, scales = "free") +
  scale_color_manual(values = lake_col_full) +
  theme(legend.position = "none",
      strip.text = element_text(size = 16),
      strip.background = element_blank()) +
  scale_y_continuous(limits = c(0, 10)) +
  ggtitle("Pollution tolerance (2020-2024)") +
  ylab("Modified HBI (family level)")

# Save plot
ggsave(here("plots", "mHBI_hhl.pdf"), width = 12, height = 8)
```

Run stats:

```{r}
lm(avg_value ~ Year, 
   data = indices %>% 
     filter(Biotic_index == "mHBI") %>% 
     filter(`Lake Name` == "Big Hawk Lake")) %>%
  summary() # coeff -0.1034, non sig
lm(avg_value ~ Year, 
   data = indices %>% 
     filter(Biotic_index == "mHBI") %>% 
     filter(`Lake Name` == "Halls Lake")) %>%
  summary() # coeff -0.14579, SIG!
lm(avg_value ~ Year, 
   data = indices %>% 
     filter(Biotic_index == "mHBI") %>% 
     filter(`Lake Name` == "Little Hawk Lake")) %>%
  summary() # coeff -0.1428, non sig
```

Get averages for executive summary:

```{r}
indices %>% 
  filter(Biotic_index == "mHBI") %>% 
  # filter(`Lake Name` == "Halls Lake") %>%
  # filter(`Lake Name` == "Big Hawk Lake") %>%
  filter(`Lake Name` == "Little Hawk Lake") %>%
  group_by(Year) %>% 
  na.omit() %>% 
  summarize(mean(avg_value))

rect_data %>% filter(Biotic_index == "mHBI")
```

Extrapolate mHBI in five years for Halls Lake.

```{r}
pred <- read_tsv(here("data", "halls_predicted.txt"))
pred <- bind_rows(indices %>% 
  filter(Biotic_index == "mHBI") %>% 
  filter(`Lake Name` == "Halls Lake"), pred)

predicted <- pred %>%
  mutate(predicted_vals = coalesce(avg_value, predict(lm(avg_value ~ Year),
                                                      across(Year))))

predicted %>% filter(Year == 2030) %>% pull(predicted_vals)
```

#### Percent EOT

```{r}
indices %>% 
  filter(Biotic_index == "Perc_EOT") %>% 
  ggplot(aes(x = Year, y = avg_value)) +
  geom_rect(data = rect_data %>% filter(Biotic_index == "Perc_EOT"), inherit.aes = FALSE,
  aes(xmin=-Inf, xmax=+Inf, ymin = ymin, ymax = ymax, fill = Category), alpha = 0.4) +
  scale_fill_manual(values = c("Fair" = "#fdae61", "Good" = "#abdda4", "Excellent" = "#3288bd")) +
  geom_point(aes(color = `Lake Name`, group = Year), size = 3) +
  geom_point(aes(group = Year), size = 3, pch = 21, color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "grey45", aes(group = 1)) +
  facet_wrap(~`Lake Name`, scales = "free") +
  scale_color_manual(values = lake_col_full) +
  theme(legend.position = "none",
      strip.text = element_text(size = 16),
      strip.background = element_blank()) +
  ggtitle("Diversity of pollution-intolerant groups (2020-2024)") +
  ylab("Percent EOT")

ggsave(here("plots", "eot_hhl.pdf"), width = 12, height = 8)
```

Run stats:

```{r}
lm(avg_value ~ Year, 
   data = indices %>% 
     filter(Biotic_index == "Perc_EOT") %>% 
     filter(`Lake Name` == "Big Hawk Lake")) %>%
  summary() # coeff -0.4659, non sig
lm(avg_value ~ Year, 
   data = indices %>% 
     filter(Biotic_index == "Perc_EOT") %>% 
     filter(`Lake Name` == "Halls Lake")) %>%
  summary() # coeff -1.493, non sig
lm(avg_value ~ Year, 
   data = indices %>% 
     filter(Biotic_index == "Perc_EOT") %>% 
     filter(`Lake Name` == "Little Hawk Lake")) %>%
  summary() # coeff -1.8733, sig @ * -> 0.05
```

### Algal growth

Import and tidy data:

```{r}
halls_veg <- read_tsv(here("data", "halls_veg.txt")) %>% 
  pivot_longer(cols = 4:15, names_to = "site_code", values_to = "value") %>% 
  group_by(Year, site_code) %>% 
  mutate(sum_algae = sum(value))

hawk_veg <- read_tsv(here("data", "hawk_veg.txt")) %>% 
  pivot_longer(cols = 4:11, names_to = "site_code", values_to = "value") %>% 
  group_by(Year, site_code) %>% 
  mutate(sum_algae = sum(value))

# Join datasets together
veg <- bind_rows(halls_veg, hawk_veg)

# Add in proper lake names
veg <-
  veg %>% 
  mutate(`Lake Name` = case_when(str_detect(site_code, regex("HALL")) ~ "Halls Lake",
                                 str_detect(site_code, regex("BHWK")) ~ "Big Hawk Lake",
                                 str_detect(site_code, regex("LHWK")) ~ "Little Hawk Lake"))
```

Build plots.

```{r}
p_algae <- veg %>% 
  ggplot(aes(x = Year, y = value, group = Vegetation_type, fill = Vegetation_type)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_grid(~`Lake Name`, scales = "free") +
  theme(strip.text = element_text(size = 10, face = "bold"),
        strip.background = element_blank(),
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.text.y = element_text(size = 9),
        axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
        legend.position = "bottom") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks = 2019:2024) +
  scale_fill_manual(values = c("#186484", "#448cbc", "#b4d6e7"),
                    name = "Type of algae") +
  # scale_fill_manual(values = c("#596b36", "#a9cc68", "#fcf79b"), 
  #                   name = "Type of algae") +
  # ggtitle("Algal growth (2019-2024)") +
  ylab("Abundance of algae")
  # ylab("Algal abundance, summed across sites")

p_algae
# Save plot for executive summary
ggsave(here("plots", "algae_hhl.pdf"), width = 3.5, height = 3) # for exec sum

# For AGM presentation
p_algae +
  theme(strip.text = element_text(size = 14, face = "bold"),
        strip.background = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 16),
        axis.title = element_text(size = 14),
        axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        legend.position = "bottom") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(breaks = 2019:2024) +
  scale_fill_manual(values = c("#186484", "#448cbc", "#b4d6e7"),
                    name = "Type of algae") +
  # scale_fill_manual(values = c("#596b36", "#a9cc68", "#fcf79b"), 
  #                   name = "Type of algae") +
  ggtitle("Algal growth (2019-2024)") +
  ylab("Abundance of algae")

ggsave(here("plots", "algae_hhl_pres.pdf"), width = 8, height = 5.5) # for pres
```

Run stats.

```{r}
lm(sum_algae ~ Year, 
   data = veg %>% 
     filter(`Lake Name` == "Big Hawk Lake")) %>%
  summary() # coeff 0.48529, SIG @ *** -> 5.23e-08
lm(sum_algae ~ Year, 
   data = veg %>% 
     filter(`Lake Name` == "Halls Lake")) %>%
  summary() # coeff 0.47095, SIG @ *** -> 3.02e-12
lm(sum_algae ~ Year, 
   data = veg %>% 
     filter(`Lake Name` == "Little Hawk Lake")) %>%
  summary() # coeff 0.20000, sig @ * -> 0.0245
```

## Lake levels

The Trent Severn Waterway monitors lake levels for the HHPLOA lakes at two gauges: one at Hawk Lake Dam and the other at Buttermilk Falls. Data for the current year are available visually [here](https://www.pc.gc.ca/apps/WaterLevels/?siteId=100419&lat=44.71941044950733&lng=-78.7716293334961&z=9&_ga=2.155594553.368853008.1747062876-2016780462.1742352049&_gl=1*1u88omb*_ga*MjAxNjc4MDQ2Mi4xNzQyMzUyMDQ5*_ga_PC690N3X7Z*czE3NDcwNjc0MjYkbzMkZzAkdDE3NDcwNjc0MjYkajAkbDAkaDA.) and for previous years through the Coalition for Equitable Water Flow [here](https://www.cewf.ca/archives/water-level-summaries).

Because raw data are not provided at the links above, we obtained the full dataset (monthly levels from 1988-2025) by contacting Trent-Severn directly.

### Lake levels

Look at lake levels for 2024.

```{r}
levels <- read_tsv(here("data", "lake_levels.txt"))
levels$Date <- lubridate::mdy(levels$Date)
levels <- levels %>% 
  mutate("Year" = year(Date),
         "DOY" = yday(Date))
# There's only a single observation for 2010 for Hawk Lake so we'll remove it
levels <- levels %>% 
  filter(!(Year == 2010 & Lake == "Hawk"))

# Calculate monthly min, max, and average across all years
summary <- levels %>% 
  group_by(Lake, DOY) %>% 
  summarize(Average = mean(`Water level (m)`),
            Maximum = max(`Water level (m)`),
            Minimum = min(`Water level (m)`))

doy_months <- read_tsv(here("data", "DOY_months.txt"))
```

Build the plot for 2024 for Hawk Lake Dam.

```{r}
ggplot() +
  # Add month bars
  geom_rect(data = doy_months %>% 
              filter(Month %in% c("Jan", "Mar", "May", "Jul", "Sep", "Nov")), 
            aes(xmin = Start, xmax = End, 
                ymin = 1.5, ymax = 5),
            fill = "grey", alpha = 0.5) +
  # Add min and max levels
  geom_ribbon(data = summary %>% 
                filter(Lake == "Hawk"),
              aes(x = DOY, ymin = Minimum, ymax = Maximum), 
              fill = "dodgerblue1", alpha = 0.5) +
  # Add average over all years
  geom_line(data = summary %>% 
              filter(Lake == "Hawk"),
            aes(x = DOY, y = Average), 
            color = "dodgerblue1", linetype = "dashed", linewidth = 1) +
  # Add 2024 data only
  geom_line(data = levels %>% 
              filter(Year == 2024) %>% 
              filter(Lake == "Hawk"), 
            aes(x = DOY, y = `Water level (m)`, group = Lake, color = Lake), 
            linewidth = 1) +
  ylab("Water level (m)") +
  xlab("Day of the Year") +
  theme(panel.grid.major.y = element_line(linewidth = 0.5, linetype = "dashed", color = "grey")) +
  ggtitle("Lake levels at Hawk Lake Dam (2024)") +
  theme(legend.position = "none") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0))

# Save plot
ggsave(here("plots", "levels_hawk.pdf"), width = 10, height = 6)
```

Build the plot for 2024 for Halls Lake Dam.

```{r}
ggplot() +
  # Add month bars
  geom_rect(data = doy_months %>% 
              filter(Month %in% c("Jan", "Mar", "May", "Jul", "Sep", "Nov")), 
            aes(xmin = Start, xmax = End, 
                ymin = 1.25, ymax = 3.1),
            fill = "grey", alpha = 0.5) +
  # Add min and max levels
  geom_ribbon(data = summary %>% 
                filter(Lake == "Halls"),
              aes(x = DOY, ymin = Minimum, ymax = Maximum), 
              fill = "dodgerblue1", alpha = 0.5) +
  # Add average over all years
  geom_line(data = summary %>% 
              filter(Lake == "Halls"),
            aes(x = DOY, y = Average), 
            color = "dodgerblue1", linetype = "dashed", linewidth = 1) +
  # Add 2024 data only
  geom_line(data = levels %>% 
              filter(Year == 2024) %>% 
              filter(Lake == "Halls"), 
            aes(x = DOY, y = `Water level (m)`, group = Lake, color = Lake), 
            linewidth = 1) +
  ylab("Water level (m)") +
  xlab("Day of the Year") +
  theme(panel.grid.major.y = element_line(linewidth = 0.5, linetype = "dashed", color = "grey")) +
  ggtitle("Lake levels at Halls Lake Dam (2024)") +
  theme(legend.position = "none") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0))

# Save plot
ggsave(here("plots", "levels_halls.pdf"), width = 10, height = 6)
```

What time of year do min/max occur?

```{r}
min_time_of_year <- 
  levels %>% 
  mutate(Month = month(Date)) %>% 
  group_by(Lake, Year) %>% 
  filter(`Water level (m)` == min(`Water level (m)`)) %>% 
  dplyr::select(Lake, Year, Month) %>% 
  distinct()
max_time_of_year <- 
  levels %>% 
  mutate(Month = month(Date)) %>% 
  group_by(Lake, Year) %>% 
  filter(`Water level (m)` == max(`Water level (m)`)) %>% 
  dplyr::select(Lake, Year, Month) %>% 
  distinct()

min_time_of_year %>% 
  ggplot(aes(x = Month)) +
  geom_bar(stat = "count", aes(fill = Lake)) +
  # scale_fill_manual(values = c("#448cbc", "#58ae50"))
  facet_grid(~Lake) +
  scale_x_continuous(breaks = 1:12)
max_time_of_year %>% 
  ggplot(aes(x = Month)) +
  geom_bar(stat = "count", aes(fill = Lake)) +
  # scale_fill_manual(values = c("#448cbc", "#58ae50"))
  facet_grid(~Lake) +
  scale_x_continuous(breaks = 1:12)
```

### Fluctuations

Amount of annual fluctuation is the maximum lake level minus the minimum lake level.

Calculate fluctuations using the levels data for each year:

```{r}
fluct <- levels %>% 
  ungroup() %>% 
  group_by(Lake, Year) %>% 
  summarize(minimum = min(`Water level (m)`),
            maximum = max(`Water level (m)`),
            fluctuation = maximum-minimum)
```

Plot the data:

```{r}
fluct$Year <- as.character(fluct$Year)

# Plot the range of values:
fluct %>% 
  ggplot() +
  geom_segment(aes(x = Year, xend = Year, y = minimum, yend = maximum, color = Lake), 
               linewidth = 10, alpha = 0.75) +
  geom_line(aes(x = Year, y = fluctuation, group = 1), color = "black") +
  geom_point(aes(x = Year, y = fluctuation), pch = 21, size = 4, 
             color = "black", fill = "white") +
  # scale_x_continuous(breaks = 2013:2024) +
  ylab("Lake level (m)") +
  ggtitle("Lake level fluctuations (2013-2024)") +
  facet_wrap(~Lake,
             nrow = 2,
             scales = "free_x",
             labeller = as_labeller(c("Hawk" = "Hawk Lake Dam", "Halls" = "Halls Lake"))) +
  scale_color_manual(values = c("Hawk" = "#58ae50", "Halls" = "#448cbc")) +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_text(size = 14),
        panel.grid.major.y = element_line(linewidth = 0.5, linetype = "dashed", color = "grey"),
        axis.text.x = element_text(angle = 45, hjust = 1))

# Save plot
ggsave(here("plots", "fluct_hhl.pdf"), width = 12, height = 8)
```

Run statistics:

```{r}
lm(fluctuation ~ Year, 
   data = fluct %>% 
     filter(Lake == "Hawk")) %>%
  summary() # coeff -0.02117, SIG @ . -> 0.0931

lm(fluctuation ~ Year, 
   data = fluct %>% 
     filter(Lake == "Halls")) %>%
  summary() # coeff -0.012222, NOT SIG

# Average fluctuation over all years
fluct %>% 
  filter(Lake == "Hawk") %>% 
  ungroup() %>% 
  summarize(mean(fluctuation))

fluct %>% 
  filter(Lake == "Halls") %>% 
  ungroup() %>% 
  summarize(mean(fluctuation))

# Min and max levels
fluct %>% 
  ungroup() %>% 
  group_by(Lake) %>% 
  summarize(min(minimum), max(maximum))
```


## Map of Ontario

The following map was downloaded using Ontario GeoHub and is a cropped version of the Ontario Land Cover Compilation v2.0, available [here](https://www.arcgis.com/home/item.html?id=7aa998fdf100434da27a41f1c637382c).

The following is how the map was created; this code is not run.

```
# map <- terra::rast(here("data", "OLCC_V2_TIFF", "OLCC_V2_TIFF.tif"))
# newcrs = "+proj=longlat +datum=WGS84"
# newmap <- terra::project(map, newcrs)
# bounds <- terra::ext(-78.861312, -78.299532, 44.650277, 45.259690)
# subset <- terra::crop(newmap, bounds)
# # Save cropped object
# terra::writeRaster(subset, here("data", "ontario_map.tif"))
```

This is a WIP and so is not run either.

```
tsw <- read_tsv(here("data", "TSW_coords.txt"))
subset <- terra::rast(here("data", "ontario_map.tif"))

pal = c("#844836", "#b07156", "#b99364", "#dac3a9", "#8fa997") # , "#6baeb0", "#376988"
palette = grDevices::colorRampPalette(pal)(100)
palette <- rev(palette)

ggplot() +
  geom_spatraster(data = subset) +
  # scale_fill_gradientn(colors = palette, na.value = NA) +
  theme_map() +
  geom_path(data = tsw %>% filter(Type == "Dam" | Type == "Lake"),
             aes(x = Longitude, y = Latitude, group = Headwater, 
                 color = Headwater), linewidth = 0.5) +
  geom_point(data = tsw %>% filter(Type == "Dam" | Type == "Lake"),
             aes(x = Longitude, y = Latitude, color = Type), size = 2) +
  geom_point(data = tsw %>% filter(Type == "City") %>% filter(Name != "Dorset"),
             aes(x = Longitude, y = Latitude), color = "white", fill = "black", pch = 21) +
  geom_text(data = tsw %>% filter(Type == "City") %>% filter(Name != "Dorset"),
             aes(x = Longitude, y = Latitude, label = Name), 
            hjust = -0.1, vjust = 0.75, color = "white", size = 3) +
  # Add 'headwater' lake labels in
  geom_label(data = tsw %>% filter(Name == "Kennisis Lake" | Name == "Percy Lake" | 
                                     Name == "Redstone Lake" | Name == "Sherborne Lake" |
                                     Name == "Little Brother Lake" | Name == "Bob Lake"),
             aes(x = Longitude, y = Latitude, label = Name, color = Name), 
             size = 3, hjust = 0.5, vjust = -0.5) +
  # Add our lake names in
  geom_label(data = tsw %>% filter(Name == "Big Hawk Lake" | Name == "Little Hawk Lake" | 
                                     Name == "Halls Lake"),
             aes(x = Longitude, y = Latitude, label = Name, color = Name), 
             size = 4, hjust = 0.5, vjust = -0.5)
```

