---
title: "Lake Survey"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(cowplot)
library(lubridate)
library(grid)
library(gridExtra)
library(interactions)
library(cowplot)
library(gtExtras)
library(MVZlibrary)
theme_set(theme_cowplot())

source(here("analysis", "functions.R"))

# Import colors
lake_cols_raw <- lake_colors(raw_data = TRUE)
data_cols_warms <- data_colors(type = "warms")
data_cols <- data_colors(type = "assorted")
lake_col_full <- lake_colors(full_names = TRUE)
```

## Background

We first want to generate plots that summarize the survey results in terms of the different categories.

Survey responses have been grouped into five categories:

1.  Nature

2.  Development

3.  Legislation

4.  Lake use

5.  Amenities

# Import data

```{r}
dat <- read_csv(here("data", "lake_plan_survey.csv"))
```

# Summary plots

Build sideways stacked bar plots for survey results, divided based on the category of question.

### Lake Use

```{r}
lake_use <- dat %>% dplyr::select(Construction_of_water_ski_courses,
                                  Fishing_and_angling,
                                  Fishing_tournaments_and_derbies,
                                  Bait_fish_release,
                                  Ice_fishing,
                                  Hunting,
                                  Snowmobiling,
                                  ATV_operation,
                                  Camping_facilities,
                                  Number_of_people_using_campsites,
                                  Pollution_from_campsites) %>% 
  mutate(respondent = 1:nrow(dat)) %>% 
  pivot_longer(cols = Construction_of_water_ski_courses:Pollution_from_campsites,
               names_to = "question", values_to = "response")

lake_use <- lake_use %>% 
  na.omit() %>% 
  group_by(question) %>% 
  mutate(category = case_when(response == 5 ~ "strongly_agree",
                              response == 4 ~ "agree",
                              response == 3 ~ "neither",
                              response == 2 ~ "disagree",
                              response == 1 ~ "strongly_disagree"))
summary <- lake_use %>% 
  group_by(question, category) %>% 
  summarize(no_respondents = n())

summary <- summary %>% 
  mutate(total_respondents = sum(no_respondents),
         prop_respondents = no_respondents/total_respondents*100) %>% 
  # group_by(question, category) %>% 
  arrange(prop_respondents)

# Overall percentage of agree/strongly agree
summary %>% 
  ungroup() %>% 
  filter(category %in% c("agree", "strongly_agree")) %>% 
  summarize(agree = sum(no_respondents),
            total = sum(total_respondents),
            percent = agree/total*100)

# Percentage of agree/strongly agree by question
summary %>% 
  group_by(question) %>% 
  filter(category %in% c("agree", "strongly_agree")) %>% 
  summarize(agree_stronglyagree = sum(no_respondents),
            total = sum(total_respondents),
            percent = agree_stronglyagree/total*100)

summary$category <- factor(summary$category, levels = c("strongly_disagree", "disagree", "neither", "agree", "strongly_agree"))
summary$question <- factor(summary$question, levels = c("Pollution_from_campsites", "Bait_fish_release", "Number_of_people_using_campsites", "Ice_fishing", "Camping_facilities", "Construction_of_water_ski_courses", "Hunting", "ATV_operation", "Fishing_tournaments_and_derbies", "Fishing_and_angling", "Snowmobiling"))
```

Build plot:

```{r}
summary %>% 
  ggplot() +
  geom_bar(stat = "identity", 
           aes(x = question, y = prop_respondents, group = category, fill = category), 
           position = "stack") +
  # scale_fill_manual(values = c("#90dcfc", "#72bede", "#54a0c0", "#186484", "#045070")) +
  scale_fill_manual(values = c("#a95e37", "#d1864b", "#ffd85b", "#448cbc", "#045070")) +
  coord_flip() +
  ylab("Proportion of respondents") +
  # xlab("Land use issue") +
  theme(axis.title.x = element_text(size = 12),
        axis.text = element_text(size = 10),
        legend.position = "bottom",
        axis.title.y = element_blank()) +
  scale_y_continuous(expand = c(0, 0))

ggsave(here("plots", "lake_use_results_gradient.pdf"), height = 3, width = 7.6)
```

### Development

```{r}
dev_dat <- dat %>% 
  # dplyr::select(Shoreline_infrastructure, Cell_towers, Backlot_development) %>% 
  dplyr::select(Condominiums,	Trailer_parks, Shipping_container_dwellings, `Floating_homes/barges`, Timeshare_dwellings, Hobby_farms, `Non-residential_commercial_development`,	Cell_towers, Backlot_development, Shoreline_infrastructure, Shoreline_minimum_setback, Shoreline_lot_minimum_frontage, Enforcement_for_upkeep_of_waste_systems) %>% 
  mutate(respondent = 1:nrow(dat)) %>% 
  pivot_longer(cols = Condominiums:Enforcement_for_upkeep_of_waste_systems,
               names_to = "question", values_to = "response")
  # pivot_longer(cols = Shoreline_infrastructure:Backlot_development,
  #              names_to = "question", values_to = "response")

dev_dat <- dev_dat %>% 
  na.omit() %>% 
  group_by(question) %>% 
  mutate(category = case_when(response == 5 ~ "strongly_agree",
                              response == 4 ~ "agree",
                              response == 3 ~ "neither",
                              response == 2 ~ "disagree",
                              response == 1 ~ "strongly_disagree"))
summary_dev <- dev_dat %>% 
  group_by(question, category) %>% 
  summarize(no_respondents = n())

summary_dev <- summary_dev %>% 
  mutate(total_respondents = sum(no_respondents),
         prop_respondents = no_respondents/total_respondents*100) %>% 
  # group_by(question, category) %>% 
  arrange(prop_respondents)

summary_dev$category <- factor(summary_dev$category, levels = c("strongly_disagree", "disagree", "neither", "agree", "strongly_agree"))
summary_dev$question <- factor(summary_dev$question,
                               levels = c("Enforcement_for_upkeep_of_waste_systems", "Shoreline_lot_minimum_frontage", "Shoreline_minimum_setback", "Shoreline_infrastructure", "Cell_towers", "Backlot_development", "Timeshare_dwellings", "Shipping_container_dwellings", "Hobby_farms", "Non-residential_commercial_development", "Floating_homes/barges", "Trailer_parks", "Condominiums"))
```

Build plot:

```{r}
summary_dev %>% 
  ggplot() +
  geom_bar(stat = "identity", 
           aes(x = question, y = prop_respondents, group = category, fill = category), 
           position = "stack") +
  scale_fill_manual(values = c("#a95e37", "#d1864b", "#ffd85b", "#448cbc", "#186484")) +
  # scale_fill_manual(values = c("#90dcfc", "#72bede", "#54a0c0", "#186484", "#045070")) +
  coord_flip() +
  ylab("Proportion of respondents (%)") +
  # xlab("Land use issue") +
  theme(axis.title.x = element_text(size = 12),
        axis.text = element_text(size = 10),
        legend.position = "bottom",
        axis.title.y = element_blank()) +
  scale_y_continuous(expand = c(0, 0))

ggsave(here("plots", "dev_results_gradient.pdf"), height = 3, width = 7.6)
```

### Nature

```{r}
nat_dat <- dat %>% 
  dplyr::select(Preserve_Natural_Beauty, Preserve_Natural_Shorelines, 
                Preserve_Unique_Habitats, Protect_Native_Species, 
                Control_Invasive_Species, Improve_Water_Quality,
                Encourage_Dark_Skies, Increase_Water_Clarity,
                Decrease_Bacterial_Levels, Reduce_Algal_Growth,
                Minimize_Water_Fluctuations) %>% 
  mutate(respondent = 1:nrow(dat)) %>% 
  pivot_longer(cols = Preserve_Natural_Beauty:Minimize_Water_Fluctuations,
               names_to = "question", values_to = "response")

nat_dat <- nat_dat %>% 
  na.omit() %>% 
  group_by(question) %>% 
  mutate(category = case_when(response == 5 ~ "strongly_agree",
                              response == 4 ~ "agree",
                              response == 3 ~ "neither",
                              response == 2 ~ "disagree",
                              response == 1 ~ "strongly_disagree"))
summary_nat <- nat_dat %>% 
  group_by(question, category) %>% 
  summarize(no_respondents = n())

summary_nat <- summary_nat %>% 
  mutate(total_respondents = sum(no_respondents),
         prop_respondents = no_respondents/total_respondents*100) %>% 
  # group_by(question, category) %>% 
  arrange(prop_respondents)

summary_nat$category <- factor(summary_nat$category, levels = c("strongly_disagree", "disagree", "neither", "agree", "strongly_agree"))
summary_nat$question <- factor(summary_nat$question,
                               levels = c("Minimize_Water_Fluctuations",
                                          "Encourage_Dark_Skies",
                                          "Increase_Water_Clarity",
                                          "Preserve_Natural_Shorelines",
                                          "Decrease_Bacterial_Levels",
                                          "Protect_Native_Species",
                                          "Reduce_Algal_Growth",
                                          "Preserve_Unique_Habitats",
                                          "Preserve_Natural_Beauty",
                                          "Control_Invasive_Species",
                                          "Improve_Water_Quality"))
```

Build plot:

```{r}
summary_nat %>% 
  ggplot() +
  geom_bar(stat = "identity", 
           aes(x = question, y = prop_respondents, group = category, fill = category), 
           position = "stack") +
  scale_fill_manual(values = c("#a95e37", "#d1864b", "#ffd85b", "#448cbc", "#186484")) +
  # scale_fill_manual(values = c("#90dcfc", "#72bede", "#54a0c0", "#186484", "#045070")) +
  coord_flip() +
  ylab("Proportion of respondents (%)") +
  theme(axis.title.x = element_text(size = 12),
        axis.text = element_text(size = 10),
        legend.position = "bottom",
        axis.title.y = element_blank()) +
  scale_y_continuous(expand = c(0, 0))

ggsave(here("plots", "nat_results_gradient.pdf"), height = 3, width = 7.6)
```

### Legislation

```{r}
soc_dat <- dat %>% 
  dplyr::select(Lower_speeds_in_open_water,
                Lower_speeds_close_to_the_shoreline,
                Watercraft_wake_height,
                Lower_speeds_for_nesting_season,
                Lower_watercraft_noise_levels,
                Safer_operation_of_watercraft,	
                Invasive_species_checks) %>% 
  mutate(respondent = 1:nrow(dat)) %>% 
  pivot_longer(cols = Lower_speeds_in_open_water:Invasive_species_checks,
               names_to = "question", values_to = "response")

soc_dat <- soc_dat %>% 
  na.omit() %>% 
  group_by(question) %>% 
  mutate(category = case_when(response == 5 ~ "strongly_agree",
                              response == 4 ~ "agree",
                              response == 3 ~ "neither",
                              response == 2 ~ "disagree",
                              response == 1 ~ "strongly_disagree"))
summary_soc <- soc_dat %>% 
  group_by(question, category) %>% 
  summarize(no_respondents = n())

summary_soc <- summary_soc %>% 
  mutate(total_respondents = sum(no_respondents),
         prop_respondents = no_respondents/total_respondents*100) %>% 
  # group_by(question, category) %>% 
  arrange(prop_respondents)

summary_soc$category <- factor(summary_soc$category, levels = c("strongly_disagree", "disagree", "neither", "agree", "strongly_agree"))
summary_soc$question <- factor(summary_soc$question,
                               levels = c("Lower_speeds_in_open_water",
                                          "Lower_watercraft_noise_levels",
                                          "Watercraft_wake_height",
                                          "Safer_operation_of_watercraft",
                                        "Lower_speeds_for_nesting_season",
                                    "Lower_speeds_close_to_the_shoreline",
                                          "Invasive_species_checks"))
```

Build plot:

```{r}
summary_soc %>% 
  ggplot() +
  geom_bar(stat = "identity", 
           aes(x = question, y = prop_respondents, group = category, fill = category), 
           position = "stack") +
  scale_fill_manual(values = c("#a95e37", "#d1864b", "#ffd85b", "#448cbc", "#186484")) +
  # scale_fill_manual(values = c("#90dcfc", "#72bede", "#54a0c0", "#186484", "#045070")) +
  coord_flip() +
  ylab("Proportion of respondents (%)") +
  theme(axis.title.x = element_text(size = 12),
        axis.text = element_text(size = 10),
        legend.position = "bottom",
        axis.title.y = element_blank()) +
  scale_y_continuous(expand = c(0, 0))

ggsave(here("plots", "soc_results_gradient.pdf"), height = 2, width = 7.6)
```

### Amenities

```{r}
ame_dat <- dat %>% 
  dplyr::select(Frequency_of_O.P.P._marine_patrols,
                Public_boat_launches,
                Availability_of_safe_hiking_trails,
                Availability_of_access_to_managed_snowmobile_trails,
                Availability_of_groomed_cross_country_ski_trails,
                Access_to_safe_ATV_trails,
                Reliable_access_to_high_speed_internet,
                Access_to_reliable_cellular_coverage,
                Reopening_of_Hawk_Lake_landfill,
                Number_of_businesses_offering_gasoline_fueling,
                `Number_and_location_of_parking_facilities_[BHL]`,
                `Number_and_location_of_parking_facilities_[LHL]`,
                Number_of_public_beaches) %>% 
  mutate(respondent = 1:nrow(dat)) %>% 
  pivot_longer(cols = Frequency_of_O.P.P._marine_patrols:Number_of_public_beaches,
               names_to = "question", values_to = "response")

ame_dat <- ame_dat %>% 
  na.omit() %>% 
  group_by(question) %>% 
  mutate(category = case_when(response == 5 ~ "strongly_agree",
                              response == 4 ~ "agree",
                              response == 3 ~ "neither",
                              response == 2 ~ "disagree",
                              response == 1 ~ "strongly_disagree"))
summary_ame <- ame_dat %>% 
  group_by(question, category) %>% 
  summarize(no_respondents = n())

summary_ame <- summary_ame %>% 
  mutate(total_respondents = sum(no_respondents),
         prop_respondents = no_respondents/total_respondents*100) %>% 
  # group_by(question, category) %>% 
  arrange(prop_respondents)

summary_ame$category <- factor(summary_ame$category, levels = c("strongly_disagree", "disagree", "neither", "agree", "strongly_agree"))
summary_ame$question <- factor(summary_ame$question,
                               levels = c("Reliable_access_to_high_speed_internet", "Access_to_reliable_cellular_coverage",
                                          "Availability_of_safe_hiking_trails",
                                          "Number_of_businesses_offering_gasoline_fueling",
                                          "Number_and_location_of_parking_facilities_[BHL]",
                "Number_and_location_of_parking_facilities_[LHL]",
                "Public_boat_launches",
                "Availability_of_groomed_cross_country_ski_trails",
                "Number_of_public_beaches",
                "Availability_of_access_to_managed_snowmobile_trails",
                "Frequency_of_O.P.P._marine_patrols",
                "Access_to_safe_ATV_trails",
                "Reopening_of_Hawk_Lake_landfill"))
```

Build plot:

```{r}
summary_ame %>% 
  ggplot() +
  geom_bar(stat = "identity", 
           aes(x = question, y = prop_respondents, group = category, fill = category), 
           position = "stack") +
  scale_fill_manual(values = c("#a95e37", "#d1864b", "#ffd85b", "#448cbc", "#186484")) +
  # scale_fill_manual(values = c("#90dcfc", "#72bede", "#54a0c0", "#186484", "#045070")) +
  coord_flip() +
  ylab("Proportion of respondents (%)") +
  theme(axis.title.x = element_text(size = 12),
        axis.text = element_text(size = 10),
        legend.position = "bottom",
        axis.title.y = element_blank()) +
  scale_y_continuous(expand = c(0, 0))

ggsave(here("plots", "ame_results_gradient.pdf"), height = 3, width = 7.6)
```

### Average value across each category

```{r}
# Retrieve question names for each category
leg_vars <- unique(soc_dat$question)
lakeuse_vars <- unique(lake_use$question)
amen_vars <- unique(ame_dat$question)
dev_vars <- unique(dev_dat$question)
nat_vars <- unique(nat_dat$question)
```

Do calculations for averages.

```{r}
### Legislation
# Calculate average values for each category from above avgs
leg_overall_mean <- soc_dat %>%
  ungroup() %>% 
  summarise(m = mean(response, na.rm = TRUE)) %>%
  pull(m)
# Calculate average values for each question
leg_question_means <- soc_dat %>%
  group_by(question) %>%
  summarise(m = mean(response, na.rm = TRUE)) %>% 
  mutate(category = "legislation",
         overall_mean = leg_overall_mean)

### Nature
# Calculate average values for each category from above avgs
nat_overall_mean <- nat_dat %>%
  ungroup() %>% 
  summarise(m = mean(response, na.rm = TRUE)) %>%
  pull(m)
# Calculate average values for each question
nat_question_means <- nat_dat %>%
  group_by(question) %>%
  summarise(m = mean(response, na.rm = TRUE)) %>% 
  mutate(category = "nature",
         overall_mean = nat_overall_mean)

### Development
# Calculate average values for each category from above avgs
dev_overall_mean <- dev_dat %>%
  ungroup() %>% 
  summarise(m = mean(response, na.rm = TRUE)) %>%
  pull(m)
# Calculate average values for each question
dev_question_means <- dev_dat %>%
  group_by(question) %>%
  summarise(m = mean(response, na.rm = TRUE)) %>% 
  mutate(category = "development",
         overall_mean = dev_overall_mean)

### Lake use
# Calculate average values for each category from above avgs
lakeuse_overall_mean <- lake_use %>%
  ungroup() %>% 
  summarise(m = mean(response, na.rm = TRUE)) %>%
  pull(m)
# Calculate average values for each question
lakeuse_question_means <- lake_use %>%
  group_by(question) %>%
  summarise(m = mean(response, na.rm = TRUE)) %>% 
  mutate(category = "lake_use",
         overall_mean = lakeuse_overall_mean)

### Amenities
# Calculate average values for each category from above avgs
amen_overall_mean <- ame_dat %>%
  ungroup() %>% 
  summarise(m = mean(response, na.rm = TRUE)) %>%
  pull(m)
# Calculate average values for each question
amen_question_means <- ame_dat %>%
  group_by(question) %>%
  summarise(m = mean(response, na.rm = TRUE)) %>% 
  mutate(category = "amenities",
         overall_mean = amen_overall_mean)

### Bind them all together
# mean_dat <- bind_rows(leg_question_means,
#                       nat_question_means,
#                       dev_question_means,
#                       lakeuse_question_means,
#                       amen_question_means)
# mean_plot_dat <- bind_rows(ame_dat %>% mutate(group = "amenities"),
#                            soc_dat %>% mutate(group = "legislation"),
#                            dev_dat %>% mutate(group = "development"),
#                            nat_dat %>% mutate(group = "nature"),
#                            lake_use %>% mutate(group = "lake_use"))
```

Build plots.

```{r}
### Legislation
soc_dat %>% 
  ggplot(aes(x = reorder(question, response), y = response)) +
  geom_boxplot(outlier.alpha = 0.3) +
  geom_point(data = leg_question_means, aes(x = question, y = m),
             color = "cornflowerblue", size = 2) +
  geom_hline(yintercept = unique(leg_question_means$overall_mean), color = "coral") +
  ggtitle("Legislation questions: response scores") +
  ylab("Response score") +
  xlab("Survey question") +
  theme(axis.text.x = element_blank(),
        # axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        axis.text.y = element_text(size = 10),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 14))
ggsave(here("plots/leg_means.pdf"), width = 6, height = 4)

### Lake use
lake_use %>% 
  ggplot(aes(x = reorder(question, response), y = response)) +
  geom_boxplot(outlier.alpha = 0.3) +
  geom_point(data = lakeuse_question_means, aes(x = question, y = m),
             color = "cornflowerblue", size = 2) +
  geom_hline(yintercept = unique(lakeuse_question_means$overall_mean), color = "coral") +
  ggtitle("Lake use questions: response scores") +
  ylab("Response score") +
  xlab("Survey question") +
  theme(axis.text.x = element_blank(),
        # axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        axis.text.y = element_text(size = 10),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 14))
ggsave(here("plots/lakeuse_means.pdf"), width = 6, height = 4)

### Development
dev_dat %>% 
  ggplot(aes(x = reorder(question, response), y = response)) +
  geom_boxplot(outlier.alpha = 0.3) +
  geom_point(data = dev_question_means, aes(x = question, y = m),
             color = "cornflowerblue", size = 2) +
  geom_hline(yintercept = unique(dev_question_means$overall_mean), color = "coral") +
  ggtitle("Development questions: response scores") +
  ylab("Response score") +
  xlab("Survey question") +
  theme(axis.text.x = element_blank(),
        # axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        axis.text.y = element_text(size = 10),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 14))
ggsave(here("plots/dev_means.pdf"), width = 6, height = 4)

### Nature
nat_dat %>% 
  ggplot(aes(x = reorder(question, response), y = response)) +
  geom_boxplot(outlier.alpha = 0.3) +
  geom_point(data = nat_question_means, aes(x = question, y = m),
             color = "cornflowerblue", size = 2) +
  geom_hline(yintercept = unique(nat_question_means$overall_mean), color = "coral") +
  ggtitle("Nature questions: response scores") +
  ylab("Response score") +
  xlab("Survey question") +
  theme(axis.text.x = element_blank(),
        # axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        axis.text.y = element_text(size = 10),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 14))
ggsave(here("plots/nature_means.pdf"), width = 6, height = 4)

### Amenities
ame_dat %>% 
  ggplot(aes(x = reorder(question, response), y = response)) +
  geom_boxplot(outlier.alpha = 0.3) +
  geom_point(data = amen_question_means, aes(x = question, y = m),
             color = "cornflowerblue", size = 2) +
  geom_hline(yintercept = unique(amen_question_means$overall_mean), color = "coral") +
  ggtitle("Amenities questions: response scores") +
  ylab("Response score") +
  xlab("Survey question") +
  theme(axis.text.x = element_blank(),
        # axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        axis.text.y = element_text(size = 10),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 14)) +
  scale_y_continuous(limits = c(1, 5))
ggsave(here("plots/amen_means.pdf"), width = 6, height = 4)
```

Save all of the above plots.

```{r}
p_nat
p_leg
p_lakeuse
p_dev
p_amen
```

# Testing for predictors

We want to see if there are any significant predictors for survey responses. To do so, we assessed survey responses based on four factors:

1.  Age of association member

2.  Length of property ownership

3.  The lake that the member has property on

4.  Type of property access

### Specify response variables (i.e., corresponding columns)

```{r}
# Nature-related questions
nature_resp <- cbind(dat$Preserve_Natural_Beauty,
                     dat$Preserve_Natural_Shorelines,
                     dat$Preserve_Unique_Habitats,
                     dat$Protect_Native_Species, 
                     dat$Control_Invasive_Species,
                     dat$Improve_Water_Quality,
                     dat$Encourage_Dark_Skies, 
                     dat$Increase_Water_Clarity, 
                     dat$Decrease_Bacterial_Levels,
                     dat$Reduce_Algal_Growth, 
                     dat$Minimize_Water_Fluctuations)

# Development-related questions
dev_resp <- cbind(dat$Condominiums,	
                  dat$Trailer_parks,
                  dat$Shipping_container_dwellings,
                  dat$`Floating_homes/barges`,
                  dat$Timeshare_dwellings, 
                  dat$Hobby_farms,
                  dat$`Non-residential_commercial_development`,	
                  dat$Cell_towers,
                  dat$Backlot_development,	
                  dat$Shoreline_infrastructure,	
                  dat$Shoreline_minimum_setback,
                  dat$Shoreline_lot_minimum_frontage,	
                  dat$Enforcement_for_upkeep_of_waste_systems)

# Legislation-related questions
leg_resp <- cbind(dat$Lower_speeds_in_open_water,
                  dat$Lower_speeds_close_to_the_shoreline,
                  dat$Watercraft_wake_height,
                  dat$Lower_speeds_for_nesting_season,
                  dat$Lower_watercraft_noise_levels,
                  dat$Safer_operation_of_watercraft,	
                  dat$Invasive_species_checks)

# Lake use-related questions
lakeuse_resp <- cbind(dat$Construction_of_water_ski_courses, 
                      dat$Fishing_and_angling,	
                      dat$Fishing_tournaments_and_derbies,	
                      dat$Bait_fish_release,	
                      dat$Ice_fishing,
                      dat$Hunting,
                      dat$Snowmobiling,
                      dat$ATV_operation,
                      dat$Camping_facilities,
                      dat$Number_of_people_using_campsites,
                      dat$Pollution_from_campsites)

# Amenities-related questions
amen_resp <- cbind(dat$Frequency_of_O.P.P._marine_patrols,
                   dat$Public_boat_launches,
                   dat$Availability_of_safe_hiking_trails,
                   dat$Availability_of_access_to_managed_snowmobile_trails,
                   dat$Availability_of_groomed_cross_country_ski_trails,
                   dat$Access_to_safe_ATV_trails,
                   dat$Reliable_access_to_high_speed_internet,
                   dat$Access_to_reliable_cellular_coverage,
                   dat$Reopening_of_Hawk_Lake_landfill,
                   dat$Number_of_businesses_offering_gasoline_fueling,
                   dat$`Number_and_location_of_parking_facilities_[BHL]`,
                   dat$`Number_and_location_of_parking_facilities_[LHL]`,
                   dat$Number_of_public_beaches)
```

### Run MANOVAs

```{r, results='hide'}
# Nature variables
nature_lake_manova <- manova(nature_resp ~ Lake, data = dat)
nature_age_manova <- manova(nature_resp ~ Age, data = dat)
nature_duration_manova <- manova(nature_resp ~ Duration, data = dat)
nature_access_manova <- manova(nature_resp ~ Access, data = dat)

# Development variables
dev_lake_manova <- manova(dev_resp ~ Lake, data = dat)
dev_age_manova <- manova(dev_resp ~ Age, data = dat)
dev_duration_manova <- manova(dev_resp ~ Duration, data = dat)
dev_access_manova <- manova(dev_resp ~ Access, data = dat)

# Legislation variables
leg_lake_manova <- manova(leg_resp ~ Lake, data = dat)
leg_age_manova <- manova(leg_resp ~ Age, data = dat)
leg_duration_manova <- manova(leg_resp ~ Duration, data = dat)
leg_access_manova <- manova(leg_resp ~ Access, data = dat)

# Lake use variables
lakeuse_lake_manova <- manova(lakeuse_resp ~ Lake, data = dat)
lakeuse_age_manova <- manova(lakeuse_resp ~ Age, data = dat)
lakeuse_duration_manova <- manova(lakeuse_resp ~ Duration, data = dat)
lakeuse_access_manova <- manova(lakeuse_resp ~ Access, data = dat)

# Amenities variables
amen_lake_manova <- manova(amen_resp ~ Lake, data = dat)
amen_age_manova <- manova(amen_resp ~ Age, data = dat)
amen_duration_manova <- manova(amen_resp ~ Duration, data = dat)
amen_access_manova <- manova(amen_resp ~ Access, data = dat)
```

### MANOVA summaries and significant predictors

```{r}
summary(nature_lake_manova)
summary(nature_access_manova)
summary(nature_age_manova)
summary(nature_duration_manova)

summary(dev_lake_manova)
summary(dev_access_manova) # sig
summary(dev_age_manova)
summary(dev_duration_manova)

summary(leg_lake_manova)
summary(leg_access_manova)
summary(leg_age_manova) # sig
summary(leg_duration_manova)

summary(lakeuse_lake_manova)
summary(lakeuse_access_manova) # sig
summary(lakeuse_age_manova) # sig
summary(lakeuse_duration_manova)

summary(amen_lake_manova) # sig
summary(amen_access_manova) # sig
summary(amen_age_manova)
summary(amen_duration_manova)
```

Look at which response variables were significant. To determine which variables are driving the significant effect, we can run separate ANOVAs for each response variable. We report MANOVA results using Pillai's Trace, which is the multivariate effect size/statistic (wherein larger values indicate stronger separation among groups), and the p-value.

```{r}
# Survey results where lake access affects response
summary(dev_access_manova) # sig PT = 0.234, p = 0.036
summary.aov(dev_access_manova)

summary(lakeuse_access_manova) # sig PT = 0.23537, p = 0.02495
summary.aov(lakeuse_access_manova)

summary(amen_access_manova) # sig PT = 0.76122, p = 0.06155
summary.aov(amen_access_manova)

# Survey results where age affects response
summary(leg_age_manova) # sig PT = 0.12044, p = 0.06702
summary.aov(leg_age_manova)
summary(lakeuse_age_manova) # sig PT = 0.23207, p = 0.02861
summary.aov(lakeuse_age_manova)

# Survey results where specific lake affects response
summary(amen_lake_manova) # sig PT = 1.6541, p = 0.002655
summary.aov(amen_lake_manova)
```

```{r}
# Re-run MANOVA for only our three lakes + Kennisis River
dat_hhlpoa <- dat %>% filter(Lake %in% c("Big Hawk Lake", "Halls Lake", "Kennisis River", "Little Hawk Lake"))
amen_resp_hhlpoa <- cbind(dat_hhlpoa$Frequency_of_O.P.P._marine_patrols,
                   dat_hhlpoa$Public_boat_launches,
                   dat_hhlpoa$Availability_of_safe_hiking_trails,
                   dat_hhlpoa$Availability_of_access_to_managed_snowmobile_trails,
                   dat_hhlpoa$Availability_of_groomed_cross_country_ski_trails,
                   dat_hhlpoa$Access_to_safe_ATV_trails,
                   dat_hhlpoa$Reliable_access_to_high_speed_internet,
                   dat_hhlpoa$Access_to_reliable_cellular_coverage,
                   dat_hhlpoa$Reopening_of_Hawk_Lake_landfill,
                   dat_hhlpoa$Number_of_businesses_offering_gasoline_fueling,
                   dat_hhlpoa$`Number_and_location_of_parking_facilities_[BHL]`,
                   dat_hhlpoa$`Number_and_location_of_parking_facilities_[LHL]`,
                   dat_hhlpoa$Number_of_public_beaches)
# Run the MANOVA
amen_lake_hhlpoa_manova <- manova(amen_resp_hhlpoa ~ Lake, data = dat_hhlpoa)
summary(amen_lake_hhlpoa_manova) # sig PT = 1.3231, p = 0.003087
```

### Build plots

#### Age as a predictor

Only build plots for significant predictors. First, let's look at categories where age was found to be a significant predictor for survey response. 

```{r}
# For age, want to extract lake use and legislation vals
# Calculate group-based mean responses
age_means <- dat %>%
  mutate(
    legislative = rowMeans(
      dplyr::select(., all_of(leg_vars)),
      na.rm = TRUE),
    lake_use = rowMeans(
      dplyr::select(., all_of(lakeuse_vars)),
      na.rm = TRUE))
age_plot_df <- age_means %>%
  dplyr::select(Age, legislative, lake_use) %>%
  pivot_longer(
    cols = c(legislative, lake_use),
    names_to = "category",
    values_to = "score"
  )
age_plot_summary <- age_plot_df %>%
  group_by(Age, category) %>%
  summarise(
    mean = mean(score, na.rm = TRUE),
    sd = sd(score, na.rm = TRUE),
    n = sum(!is.na(score)),
    se = sd / sqrt(n),
    .groups = "drop"
  )
age_plot_summary$Age <- factor(age_plot_summary$Age, levels = c("Under 25 years old", "26 - 55 years old", "56 + years old"))
```

Build the plot for age:

```{r}
age_plot_summary %>% 
  ggplot(aes(x = Age, y = mean, color = Age)) +
  geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                width = 0.35,
                linewidth = 1) +
  geom_point(size = 2) +
  ylab("Respondent agreement") +
  xlab("Age group") +
  scale_color_manual(values = c("#208c74", "#a82c94", "#bd724b")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        legend.position = "none") +
  facet_wrap(~category, scales = "free_y") +
  scale_y_continuous(limits = c(0, 5))

ggsave(here("plots/age_means_plot.pdf"), width = 6, height = 4)
```

#### Access as a predictor

```{r}
# For access, want to extract lake use, development, and amenities vals
# Calculate group-based mean responses
access_means <- dat %>%
  mutate(
    development = rowMeans(
      dplyr::select(., all_of(dev_vars)),
      na.rm = TRUE),
    lake_use = rowMeans(
      dplyr::select(., all_of(lakeuse_vars)),
      na.rm = TRUE),
    amen = rowMeans(
      dplyr::select(., all_of(amen_vars)),
      na.rm = TRUE))
access_plot_df <- access_means %>%
  dplyr::select(Access, development, lake_use, amen) %>%
  pivot_longer(
    cols = c(development, lake_use, amen),
    names_to = "category",
    values_to = "score"
  )
access_plot_summary <- access_plot_df %>%
  group_by(Access, category) %>%
  summarise(
    mean = mean(score, na.rm = TRUE),
    sd = sd(score, na.rm = TRUE),
    n = sum(!is.na(score)),
    se = sd / sqrt(n),
    .groups = "drop"
  )
access_plot_summary$Access <- factor(access_plot_summary$Access, levels = c("Private road", "Public road", "Water"))
```

Build the plot for access:

```{r}
access_plot_summary %>% 
  ggplot(aes(x = Access, y = mean, color = Access)) +
  geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                width = 0.35,
                linewidth = 1) +
  geom_point(size = 2) +
  ylab("Respondent agreement") +
  xlab("Property access type") +
  scale_color_manual(values = c("#f6c34c", "#6b4e20", "#306281")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        legend.position = "none") +
  facet_wrap(~category, scales = "free_y") +
  scale_y_continuous(limits = c(0, 5))

ggsave(here("plots/access_means_plot.pdf"), width = 9, height = 4)
```

#### Waterbody as a predictor

```{r}
# For waterbody, want to extract amenities vals
# Calculate group-based mean responses
waterbody_means <- dat %>%
  mutate(
    amen = rowMeans(
      dplyr::select(., all_of(amen_vars)),
      na.rm = TRUE))
waterbody_plot_df <- waterbody_means %>%
  dplyr::select(Lake, amen) %>%
  pivot_longer(
    cols = c(amen),
    names_to = "category",
    values_to = "score"
  )
waterbody_plot_summary <- waterbody_plot_df %>%
  group_by(Lake, category) %>%
  summarise(
    mean = mean(score, na.rm = TRUE),
    sd = sd(score, na.rm = TRUE),
    n = sum(!is.na(score)),
    se = sd / sqrt(n),
    .groups = "drop"
  )
# waterbody_plot_summary$Lake <- factor(waterbody_plot_summary$Lake, levels = c(TODO))
```

Build the plot for waterbody:

```{r}
waterbody_plot_summary %>% 
  filter(Lake %in% c("Big Hawk Lake", "Little Hawk Lake", "Kennisis River", "Halls Lake")) %>% 
  ggplot(aes(x = Lake, y = mean, color = Lake)) +
  geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                width = 0.35,
                linewidth = 1) +
  geom_point(size = 2) +
  ylab("Respondent agreement") +
  xlab("Waterbody where property is located") +
  scale_color_manual(values = lake_col_full) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        legend.position = "none") +
  facet_wrap(~category, scales = "free_y") +
  scale_y_continuous(limits = c(0, 5))

ggsave(here("plots/waterbody_means_plot.pdf"), width = 3, height = 4)
```

### Build tables

```{r}
results <- tibble(
  Response = paste("Response", 1:13),
  F_value = c(6.4713, 5.0354, 1.3434, 4.0149, 4.3188,
              4.4568, 5.2688, 0.9747, 3.349,
              0.9987, 2.2859, 0.2315, 0.6872),
  p_value = c(0.001964, 0.007526, 0.2638, 0.01981,
              0.01483, 0.01301, 0.006041,
              0.3794, 0.03749, 0.3706,
              0.1049, 0.7936, 0.5044)
)

results <- results %>%
  mutate(sig = case_when(
    p_value < 0.001 ~ "***",
    p_value < 0.01  ~ "**",
    p_value < 0.05  ~ "*",
    TRUE ~ "ns"
  ))
results <- results %>%
  arrange(p_value) %>%
  mutate(Response = factor(Response, levels = Response))

tbl <- results %>%
  filter(p_value <= 0.05) %>% 
  gt::gt() %>%
  gtExtras::gt_hulk_col_numeric(p_value, trim = TRUE, reverse = TRUE)
```

## Top 11 issues

Import and process data.

```{r}
top_issues <- read_tsv(here("data", "survey_top_issues.txt")) %>% 
  pivot_longer(cols = "Top_11_enviro_issues":"...10", names_to = "tmp", values_to = "answer_code") %>% 
  dplyr::select(-tmp) %>% 
  na.omit() %>% 
  mutate(issue = case_when(answer_code == "1a" ~ "natural_beauty",
                           answer_code == "1b" ~ "natural_shoreline",
                           answer_code == "1c" ~ "unique_habitats",
                           answer_code == "1d" ~ "native_species",
                           answer_code == "1e" ~ "invasive_species",
                           answer_code == "1f" ~ "water_quality",
                           answer_code == "1g" ~ "dark_skies",
                           answer_code == "1h" ~ "causeway",
                           answer_code == "2a" ~ "water_clarity",
                           answer_code == "2b" ~ "bacteria",
                           answer_code == "2c" ~ "algae",
                           answer_code == "2d" ~ "water_levels"))
```

Top three issues:

```{r}
# Overall
top_issues %>% 
  group_by(issue) %>% 
  summarise(n()) %>% 
  top_n(3) %>% 
  arrange(desc(`n()`))

# By lake/river
top_issues %>% 
  filter(Lake %in% c("Halls Lake", "Little Hawk Lake", "Big Hawk Lake", "Kennisis River")) %>% 
  group_by(Lake, issue) %>% 
  summarise(n()) %>% 
  top_n(3) %>% 
  arrange(desc(`n()`))
```

### Plot data

```{r}
top_issues %>% 
  count(issue, sort = TRUE) %>%
  mutate(issue = fct_reorder(issue, desc(n))) %>%
  ggplot(aes(x = issue, y = n)) +
  geom_col(fill = "cornflowerblue") +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Number of votes") +
  xlab("Issue")

ggsave(here("plots", "top_issues.pdf"), height = 4, width = 4)
```

Break up results based on the lake:

```{r}
top_issues %>% 
  filter(Lake %in% c("Halls Lake", "Little Hawk Lake", "Big Hawk Lake", "Kennisis River")) %>% 
  count(Lake, issue, sort = TRUE) %>%
  mutate(issue = fct_reorder(issue, n, .desc = TRUE)) %>%
  ggplot(aes(x = issue, y = n, fill = Lake)) +
  geom_col() +
  facet_wrap(~Lake, nrow = 2) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Number of votes") +
  xlab("Issue") +
  scale_fill_manual(values = lake_col_full) +
  theme(legend.position = "none")

ggsave(here("plots", "top_issues_bywaterbody.pdf"), height = 8, width = 8)
```

## Residency pie charts

Import data:
```{r}
res <- read_tsv(here("data/residency_survey.txt"))
res$Response <- factor(res$Response, 
                          levels = c("7+_mos", "4-6_mos", "2-3_mos", "15-30_d", "1-14d", "0_mos", "Undecided", "No", "Yes", "Already_is"))
```

Build pie charts:
```{r}
ggplot(res %>% filter(Category == "Residence_timing"), 
       aes(x = "", y = Percentage, fill = Response)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  # scale_fill_manual(values = MVZ_palette("LifeHistories")) +
  scale_fill_manual(values = MVZ_palette("WesternBirds")) +
  # scale_fill_manual(values = data_cols) +
  theme(axis.title = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom")

ggsave(here("plots/residence_timing.pdf"), width = 4.75, height = 5)

ggplot(res %>% filter(Category == "Future_residence"), 
       aes(x = "", y = Percentage, fill = Response)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  # scale_fill_manual(values = MVZ_palette("LifeHistories")) +
  scale_fill_manual(values = MVZ_palette("WesternBirds")) +
  # scale_fill_manual(values = data_cols) +
  theme(axis.title = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "bottom")

ggsave(here("plots/fulltime_residence.pdf"), width = 4.75, height = 5)
```
